\section*{Acknowledgements}
\begin{flushleft}

I want to thank everyone who supported me during my master thesis.
\vspace{1cm}

First of all, I am very grateful to Prof. Johannes Müller for making sure I could work on a topic that I am passionate about. Thank you for offering just the right amount of guidance whenever I needed it while allowing me to learn how to work independently at the same time.
\vspace{0.75cm}
Further, I would like to thank PD Dr. Mark-Oliver Rödel for agreeing to evaluate this thesis and helpful literature suggestions.
\vspace{0.75cm}
I wish to acknowledge the help I received from Dr. Catalina Pimiento, who kindly offered help with the analyses, readily shared her R-Scripts with me and was always open for questions. Thank you very much for your support!
\vspace{0.75cm}
I am thankful to Frank Tillack for access to and help with handling and measuring collection material.
\vspace{0.75cm}

I would also like to thank Irena Rostalski and Roberto Rozzi for assistance with foreign literature.

\vspace{0.75cm}
Special thanks goes to Sonja Rothkugel, who was always helpful with R and advice on statistics and method in general. Further, thank you for reading parts of this thesis in advance and giving constructive feedback.
\vspace{0.75cm}
I also want to thank the following for critically reading a first draft of this thesis: Phillip Pinder, Viviane Kremling, Aaron Czycholl and Falk Mielke. Your help is very much appreciated!
\vspace{0.75cm}
Reading first draft: Phillip, Aaron, Falk, Vivi, Maria
\vspace{0.75cm}
Lastly, I want to thank my parents and family for their unwavering support during my studies!


\end{flushleft}{