---
title: "MAthesis"
#author: "Julia Joos"
#date: "14 August 2017"
output: 
  pdf_document:
   # classoption: landscape
    toc: true
   # toc_depth: 2
   # number_sections: true
    fig_caption: true
    df_print: kable
   # fontsize: 12pt
   # geometry: margin=1in
   # documentclass: scrartcl
   # mainfont: Helvetica
  # citation_package: natbib
    keep_tex: true
  

header-includes:
    - \usepackage{setspace}
    - \doublespacing
#     
# includes:
#       before_body: titlepage.tex
    
---

```{r "setup", include=FALSE}
require("knitr")
#opts_knit$set(root.dir = "//naturkundemuseum-berlin.de/MuseumDFSRoot/Benutzer/Julia.Joos/Eigene Dateien/MA")
## on laptop: 
opts_knit$set(root.dir = "C:/Users/Jule/Documents/Uni/MA")
library(dplyr)
library(ggplot2)
library(paleoTS)
library(stats)
library(moments)
```




```{r "import data set", echo=FALSE}
#### Data basis - import data set from Excel ####
setwd("C:/Users/Jule/Documents/Uni/MA")
tidyCL<-read.csv("tortoises_tidy.csv", sep=";", header=TRUE)

#####prepare for analysis (fix column names in .csv-file after converting table from excel####
colnames(tidyCL)[6] <- "MAmin"
colnames(tidyCL)[7] <- "Mamax"
colnames(tidyCL)[17] <- "CL"
colnames(tidyCL)[18] <- "PL"
colnames(tidyCL)[21] <- "estimated"

tidyCL <-  tidyCL %>%
  mutate(Age= ((MAmin)+(as.numeric(Mamax)))/2) # %>%
  # filter(estimated=="m") %>%
  # filter(!is.na(CL))

####### import extant data ####
extant <- read.csv("MFN_testudinidae.csv", sep=";", header=TRUE, dec=".", na.strings = "NA", stringsAsFactors=FALSE)  # file: MFN_testudinidae.csv

colnames(extant)[5] <- "SCL"
colnames(extant)[11] <- "PL"
colnames(extant)[12] <- "PLmid"



Extant <- extant %>%
  mutate(CL = SCL * 10, PL=PL*10, PLmid=PLmid*10) 
```


```{r "estimating CL from PL", echo=FALSE, results='asis'}
CLPLtidy <- tidyCL %>%
  filter(!is.na(CL) & !is.na(PL)) %>%
  dplyr::select(Taxon, CL, PL, size, Age, Island, Continent, Genus) %>%
  mutate(ratio=CL/PL) %>%
  group_by(Taxon) %>% #to show ratios per Taxon, leave out to get a total ratio
  summarise(meanRatio=round(mean(ratio),2), sdRatio=round(sd(ratio),2), n=n(), min=min(ratio), max=max(ratio))

CLPLextant <- Extant %>%
  filter(!is.na(CL) & !is.na(PL)) %>%
  dplyr::select(Taxon=Species, CL, PL, PLmid, Island, Continent, Genus) %>%
  mutate(ratio=CL/PL, ratioMid=CL/PLmid)
  #group_by(Taxon) %>% #to show ratios per Taxon, leave out to get a total ratio
  #  summarise(meanRatio=round(mean(ratio),2), sdRatio=round(sd(ratio),2), n=n(), min=min(ratio), max=max(ratio))

Ratio <- CLPLextant %>%
  summarise(meanRatio=round(mean(ratio),2), sdRatio=round(sd(ratio),2), n=n(), min=min(ratio), max=max(ratio))

RatioSpecies <- CLPLextant %>%
  group_by(Taxon) %>% #to show ratios per Taxon, leave out to get a total ratio
  summarise(meanRatio=round(mean(ratio),2), sdRatio=round(sd(ratio),2), n=n(), min=min(ratio), max=max(ratio))


fossil <- tidyCL %>%
  mutate(Taxon=as.character(Taxon), Island=as.character(Island), Continent=as.character(Continent), estimated=as.character(estimated), Genus=as.character(Genus)) %>%
  dplyr::select(Taxon, CL, PL, size, estimated, Age, Island, Continent, Genus)
  
modern <- Extant %>%
  dplyr::select(Taxon=Species, CL, PL, estimated, Age, Island, Continent, Genus)

All <- bind_rows(modern, fossil)

testRatio <- All %>% #tidyCL
  dplyr::select(Taxon, CL, PL, size, estimated, Age, Island, Continent, Genus) %>%
  mutate(extraCL = PL*Ratio$meanRatio) %>%
  dplyr::select(Taxon, CL, extraCL, PL, size, estimated, Age, Island, Continent, Genus)

testRatio$CL[is.na(testRatio$CL)] <- testRatio$extraCL[is.na(testRatio$CL)]
```




\newpage
# Time Bins with sample sizes

```{r "Bin data", echo=FALSE}

PleiPlio <- testRatio #%>% #testRatio or tidyCL 
#  filter(Age < 11.000)

# PleiPlio$bin <- cut(PleiPlio$Age, c(0, 0.000001, 0.0117, 0.126, 0.781, 2.588, 3.6, 5.332, 11.608))
# 
# EpochBins <- as.vector(c("Modern", "Modern", "Upper Pleistocene", "Middle Pleistocene", "Lower Pleistocene", "Upper Pliocene", "Lower Pliocene", "Upper Miocene"))
# 
# 
# MeanBins <- as.vector(c((0+0.000001)/2, (0+0.0117)/2, (0.0117+0.126)/2, (0.126+0.781)/2, (0.781+2.588)/2,(2.588+3.6)/2, (3.6+5.332)/2, (5.332+11.608)/2))


PleiPlio$bin <- cut(PleiPlio$Age, c(0, 0.0117, 0.126, 0.781, 2.588, 3.6, 5.332, 11.608, 15.97, 23.03, 50.000))

EpochBins <- as.vector(c("Modern", "Upper Pleistocene", "Middle Pleistocene", "Lower Pleistocene", "Upper Pliocene", "Lower Pliocene", "Upper Miocene", "Middle Miocene", "Lower Miocene", "Oligocene and Eocene"))


MeanBins <- as.vector(c((0+0.0117)/2, (0.0117+0.126)/2, (0.126+0.781)/2, (0.781+2.588)/2,(2.588+3.6)/2, (3.6+5.332)/2, (5.332+11.608)/2, (11.608+15.97)/2, (15.97+23.03)/2, (23.03+50.000)/2))




Bins <- PleiPlio %>%
  select(bin) %>%
  group_by(bin) %>%
  summarise(nIndividuals=n())

BinsSpecies <- PleiPlio %>%
  select(bin, Taxon) %>%
  group_by(bin, Taxon) %>%
  summarise(nSpecies=n()) %>%
  summarise(nSpecies=n())


BinsGenera <- PleiPlio %>%
  select(bin, Genus) %>%
  group_by(bin, Genus) %>%
  summarise(nGenera=n()) %>%
  summarise(nGenera=n())

bin <- as.vector(unique(Bins$bin))



BINS <- data.frame(bin, EpochBins, MeanBins) %>%
  merge(Bins) %>%
  merge(BinsSpecies) %>%
  merge(BinsGenera) %>%
  arrange(MeanBins)

BINS$EpochBins = factor(BINS$EpochBins, levels=c("Modern", "Upper Pleistocene",
                                                             "Middle Pleistocene", "Lower Pleistocene",
                                                             "Upper Pliocene", "Lower Pliocene",
                                                             "Upper Miocene", "Middle Miocene", 
                                                              "Lower Miocene", "Oligocene and Eocene"))

#BINS <- read.table("timebins.txt", sep="\t", header=TRUE)

#kable(Bins, caption="Time bins with corresponding sample sizes (individuals)")

kable(BINS, caption="Time bins with age range, epoch name, mean age and corresponding sample sizes (on individual, species and genus level)")

PleiPlioCL <- PleiPlio %>%
  merge(BINS) %>%
  filter(EpochBins != "Oligocene and Eocene")

# SUM <- PleiPlioCL %>%
#   group_by(MeanBins) %>%
#   summarise(n=n())


  
```

```{r "Get overview over data set", echo=FALSE, fig.cap="Scatterplot of CL over time, indicating insular (triangle) and continental (circles) and colour indicating continents. Lines indicte bins, dashed line = new bins."}
scatter <- testRatio %>%
  filter(!is.na(CL)) %>%
  filter(Age < 23.000) %>%
  select(CL, Age, Island, Continent, estimated) %>%
  ggplot(aes(Age, CL)) + geom_jitter(aes(shape=Island, colour= Continent)) +
  theme_classic() +
  geom_vline(xintercept= c(0, 0.0117, 0.126, 0.781, 2.588, 3.6, 5.332, 11.608, 15.97, 23.03)) +
  geom_vline(xintercept= c(1.806, 4.466, 7.424, 9.516, 13.789, 18), linetype="dashed")


scatter

length(scatter$CL)


```

\newpage
# Smaller time bins

```{r "Bin data, smaller bins", echo=FALSE}
Miocene <- testRatio %>% #testRatio or tidyCL 
  filter(Age < 23.000)

# PleiPlio$bin <- cut(PleiPlio$Age, c(0, 0.000001, 0.0117, 0.126, 0.781, 2.588, 3.6, 5.332, 11.608))
# 
# EpochBins <- as.vector(c("Modern", "Modern", "Upper Pleistocene", "Middle Pleistocene", "Lower Pleistocene", "Upper Pliocene", "Lower Pliocene", "Upper Miocene"))
# 
# 
# MeanBins <- as.vector(c((0+0.000001)/2, (0+0.0117)/2, (0.0117+0.126)/2, (0.126+0.781)/2, (0.781+2.588)/2,(2.588+3.6)/2, (3.6+5.332)/2, (5.332+11.608)/2))


# new cuts: 1.806, 4.466, 7.424, 9.516, 13.789, 18

Miocene$bin <- cut(Miocene$Age, c(0, 0.0117, 0.126, 0.781, 1.806, 2.588, 3.6, 4.466,
                                  5.332,7.424, 9.516, 11.608, 13.789, 15.97, 18, 23.03))




EpochBins <- as.vector(c("Modern", "Upper Pleistocene", "Middle Pleistocene", "Lower Pleistocene", "Gelasian(LowPleio2)", "Upper Pliocene", "Lower Pliocene 1", "Lower Pliocene 2","Upper Miocene 1", 
                         "Upper Miocene 2","Upper Miocene 3","Middle Miocene 1","Middle Miocene 2",
                         "Lower Miocene 1", "Lower Miocene 2"))


MeanBins <- as.vector(c((0+0.0117)/2, (0.0117+0.126)/2, (0.126+0.781)/2, (0.781+1.806)/2, (1.806+2.588)/2,(2.588+3.6)/2, (3.6+4.446)/2, (4.446+5.332)/2, (5.332+7.424)/2, (7.424+9.516)/2,
                        (9.516+11.608)/2, (11.608+13.789)/2, (13.789+15.97)/2, (15.97+18)/2, 
                        (18+23.03)/2))

#1.806, 4.466, 7.424, 9.516, 13.789, 18


BinsMio <- Miocene %>%
  select(bin) %>%
  group_by(bin) %>%
  summarise(nIndividuals=n())

BinsSpeciesMio <- Miocene %>%
  select(bin, Taxon) %>%
  group_by(bin, Taxon) %>%
  summarise(nSpecies=n()) %>%
  summarise(nSpecies=n())


BinsGeneraMio <- Miocene %>%
  select(bin, Genus) %>%
  group_by(bin, Genus) %>%
  summarise(nGenera=n()) %>%
  summarise(nGenera=n())

bin <- as.vector(unique(BinsMio$bin))



BINSMio <- data.frame(bin, EpochBins, MeanBins) %>%
  merge(BinsMio) %>%
  merge(BinsSpeciesMio) %>%
  merge(BinsGeneraMio) %>%
  arrange(MeanBins)

 BINSMio$EpochBins = factor(BINSMio$EpochBins, 
                           levels=c("Modern", "Upper Pleistocene", "Middle Pleistocene",
                                    "Lower Pleistocene", "Gelasian(LowPleio2)", "Upper Pliocene",
                                    "Lower Pliocene 1", "Lower Pliocene 2","Upper Miocene 1",
                                    "Upper Miocene 2", "Upper Miocene 3","Middle Miocene 1", 
                                    "Middle Miocene 2","Lower Miocene 1", "Lower Miocene 2"))

#BINS <- read.table("timebins.txt", sep="\t", header=TRUE)

#kable(Bins, caption="Time bins with corresponding sample sizes (individuals)")

kable(BINSMio, caption="Smaller time bins with age range, epoch name, mean age and corresponding sample sizes (on individual, species and genus level)")

PleiPlioCL <- Miocene %>%
  merge(BINSMio) 

# SUM <- PleiPlioCL %>%
#   group_by(MeanBins) %>%
#   summarise(n=n())

na <- PleiPlioCL$EpochBins[!complete.cases(PleiPlioCL$EpochBins)]


  
```


\newpage

# Maps
## fossil occurences of testudinidae

```{r "Map fossil occurrences", echo=FALSE, fig.cap="Map displaying all fossil occurrences of testudinids, with color indicating whether relevant literature was available (black if not) and if it was, whether body size data was available or not (yes and no, respectively)."}

CL <- tidyCL %>%
#  filter(Age < 11.000) %>%
  dplyr::select(Locality, Genus, Taxon, Latitude, Longitude, Country, Age) %>%
  group_by(Locality) %>%
  mutate(count= n())


FossilOccurrences<-read.csv("tortoises_occurrences.csv", sep=";", header=TRUE)

FossOcc <- FossilOccurrences %>%
  mutate(Age= (as.numeric(as.character(MA.min)) + as.numeric(as.character(Ma.max)))/2) %>%
  select(Locality, Country, Latitude, Longitude, Age, Genus, Taxon, Clavailability) %>%
#  merge(tidyCL) %>%
  group_by(Locality) %>%
  mutate(count= n(), Longitude=as.numeric(as.character(Longitude)))


#na <- FossOcc[!complete.cases(FossOcc),]

# OccMap <-  merge(Map, FossOcc, by="Locality", incomparables=TRUE) %>%
#  distinct()
  
# unique(CL$Locality) #wie viele localities
#length(unique(FossOcc$Locality[which(FossOcc$Clavailability == "yes")]))


mapWorld <- borders("world", colour="azure3", fill="azure3") # create a layer of borders, run line at the beginning (before loading plotly)

cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")


mapOc <- FossOcc %>%
  ggplot(aes(Longitude, Latitude)) +# borders("world", ylim = c(-60, 90)) +
  mapWorld +
  theme_classic() +
  geom_point(data=FossOcc, (aes(as.numeric(Longitude), Latitude, colour=FossOcc$Clavailability))) +
  geom_point(data=CL,(aes(Longitude, Latitude, colour="yes"))) +
  scale_colour_manual(values=cbbPalette)

#unique(FossOcc$CL)

mapOc

# require(plotly)
# ggplotly(mapOc)




# length(CL$Locality)
# length(unique(CL$Locality)) #wie viele localities
# length(unique(FossOcc$Locality[which(FossOcc$Clavailability == "yes")]))



```

\newpage
## body size of testudinidae

```{r "Map body size data set", echo=FALSE, fig.cap="Map displaying all localities for which body size data for testudinids was available in the literature. Size of points denotes sample size, color denotes approximate age."}


tidyCL$bin <- cut(tidyCL$Age, c(0, 0.000001, 0.0117, 0.126, 0.781, 2.588, 3.6, 5.332, 11.608, 15.97, 23.03, 50.000))



MapCL <- tidyCL %>%
  merge(BINS) %>%
#  filter(Age < 23.000) %>%
#  filter(Latitude != "-") %>%
  dplyr::select(Genus, Taxon, Latitude, Longitude, Country, CL, PL, MeanBins, size) %>%
  group_by(Latitude) %>%
  mutate(count= n()) 


mapWorld <- borders("world", colour="azure3", fill="azure3") # create a layer of borders, run line at the beginning (before loading plotly)


mapCL <- MapCL %>%
  ggplot(aes(Longitude, Latitude)) +# borders("world", ylim = c(-60, 90)) +
  mapWorld +
  theme_classic() +
  geom_point(aes(Longitude, Latitude,size=count, colour=MeanBins)) +  #colour=size, 
  scale_colour_gradientn(colors=c("orange", "red", "purple", "blue", "green"))



mapCL


```




```{r "Get overview over sample sizes", echo=FALSE, include=FALSE}
OverviewSpecies <- PleiPlioCL %>%
  filter(EpochBins != "Modern") %>%
  group_by(EpochBins, Taxon) %>%
  summarise(n=n())

OverviewSpeciesFossil <- PleiPlioCL %>%
  filter(EpochBins != "Modern") %>%
  group_by(Taxon) %>%
  summarise(n=n())

#write.table(OverviewSpecies,file="OverviewSpeciesSampleSize.txt", sep="\t", row.names = FALSE) 


OverviewGeneraTime <- PleiPlioCL %>%
  group_by(EpochBins, Genus) %>%
  summarise(n=n())

OverviewGenera <- PleiPlioCL %>%
  group_by(Genus) %>%
  summarise(n=n())

```

\newpage

# Sampling Accumulation Curve

```{r "Species Accumulation Curve", echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Species Accumulation Curve of fossil species per Locality and reference", include=FALSE}

vegan <- tidyCL %>%
  dplyr::select(Locality, Taxon) %>%
  group_by(Locality, Taxon) %>%
  summarise(n=n()) %>%
  tidyr::spread(Taxon, n, fill=0)

library(vegan)

#par(mfrow = c(2, 1))

vegan=vegan[,-1]
vegansp=specaccum(vegan,method="rarefaction", permutations=1000)
plot(vegansp,xlab="Localities",ylab="Richness", xvar="individuals", ci.type="line", ci.lty=2, ci.col="grey", col="deepskyblue4", lwd=2, main="Fossil species, CL, per Locality" )



vegan <- tidyCL %>%
  dplyr::select(Reference, Taxon) %>%
  group_by(Reference, Taxon) %>%
  summarise(n=n()) %>%
  tidyr::spread(Taxon, n, fill=0)

library(vegan)

vegan=vegan[,-1]
vegansp=specaccum(vegan,method="rarefaction", permutations=1000)
plot(vegansp,xlab="References",ylab="Richness", xvar="individuals", ci.type="line", ci.lty=2, ci.col="grey", col="deepskyblue4", lwd=2, main="Fossil species, CL, per Reference" )


```



```{r "Species Accumulation Curve with Genera", echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Sampling Accumulation Curve of fossil genera per reference"}

vegan <- tidyCL %>%
  dplyr::select(Locality, Genus) %>%
  group_by(Locality, Genus) %>%
  summarise(n=n()) %>%
  tidyr::spread(Genus, n, fill=0)




#par(mfrow = c(2, 1))


vegan=vegan[,-1]
vegansp=specaccum(vegan,method="rarefaction", permutations=1000)
# plot(vegansp,xlab="Localities",ylab="Richness", xvar="individuals", ci.type="line", ci.lty=2, ci.col="grey", col="deepskyblue4", lwd=2, main="Fossil genera, CL, per Locality" )
# --> appendix!


vegan <- tidyCL %>%
  dplyr::select(Reference, Genus) %>%
  group_by(Reference, Genus) %>%
  summarise(n=n()) %>%
  tidyr::spread(Genus, n, fill=0)

library(vegan)

vegan=vegan[,-1]
vegansp=specaccum(vegan,method="rarefaction", permutations=1000)
plot(vegansp,xlab="References",ylab="Richness", xvar="individuals", ci.type="line", ci.lty=2, ci.col="grey", col="deepskyblue4", lwd=2, main="Fossil genera, CL, per Reference")


```
\newpage

## Eurasia

```{r "Species Accumulation Curve with Genera, Eurasia", echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Sampling Accumulation Curve of fossil genera per reference, Eurasia"}

Continent <- as.vector(unique(PleiPlioCL$Continent))
Con <- as.vector(c("Asia", "Africa", "Europe", "America", "America", "America", "Europe"))#
Continents <- data.frame(Continent, Con)

overviewVegan <- PleiPlioCL %>%
  merge(Continents) %>%
  filter(EpochBins != "Modern") %>%
  group_by( Con) %>%  #Continent,
 # filter(Con=="Europe" | Con=="Asia" )%>%
  filter(!is.na(CL)) %>%
  summarise(meanCL=mean(CL), sdCL= sd(CL), n=n(), meanAge=mean(Age))


vegan <- tidyCL %>%
  merge(Continents) %>%
  filter(Con=="Europe" | Con=="Asia") %>%
  dplyr::select(Reference, Genus) %>%
  group_by(Reference, Genus) %>%
  summarise(n=n()) %>%
  tidyr::spread(Genus, n, fill=0)

#library(vegan)

vegan=vegan[,-1]
vegansp=specaccum(vegan,method="rarefaction", permutations=1000)
plot(vegansp,xlab="References",ylab="Richness", xvar="individuals", ci.type="line", ci.lty=2, ci.col="grey", col="deepskyblue4", lwd=2, main="Fossil genera, CL, per Reference")


```



\newpage

# Histograms
## all

```{r "Histograms of body size data, all", echo=FALSE, fig.cap="Distribution of body size data, logtransformed, all data."}


HistCL <- PleiPlioCL %>%
  filter( !is.na(CL)) %>%
  mutate(CL=log(CL)) %>%
#  filter(EpochBins != "Modern") %>%
  ggplot(aes(CL)) + 
  #geom_histogram( col="black", fill="gray") + 
  geom_histogram( aes(y=..density..),col="black", fill="gray")+
  #geom_density() +
  geom_density(alpha=.2, fill="darkslategray1") +
  #facet_wrap(~EpochBins, scales="free_x")+   #binwidth=10,
  theme_classic()+ 
  #theme_gray() +
  theme(panel.border = element_rect(colour = "black", fill=NA))

#HistCLModern
HistCL  






StatsAll <- PleiPlioCL %>%
  filter(!is.na(CL)) %>%
#  group_by(Island) %>%
  summarise(nCL=length(CL), #range=range(CL),
            min=min(CL), max=max(CL), 
            var=var(CL), mean=round(mean(CL),1), logm=round(mean(log10(CL)),1),
            med=round(median(CL),1), logmed=round(median(log10(CL)),1), #CLmode=mode(CL), CLlogmode=mode(log10(CL)),
            skew=round(skewness(CL),2), logsk=round(skewness(log10(CL)),2),
            kurt=round(kurtosis(CL),2), logku=round(kurtosis(log10(CL)),2) ) %>%
  unique()





```
```{r "normal distribution, echo=FALSE"}

qqnorm(PleiPlioCL$CL); qqline(PleiPlioCL$CL, col=2)
qqnorm(log10(PleiPlioCL$CL)); qqline(log10(PleiPlioCL$CL), col=2)


```



\newpage

## per time bin

```{r "Histograms of body size data, per time bin", echo=FALSE, fig.cap="Distribution of body size data per time bin, logtransformed."}


# PleiPlioCL$EpochBins = factor(PleiPlioCL$EpochBins, 
#                            levels=c("Modern", "Upper Pleistocene", "Middle Pleistocene",
#                                     "Lower Pleistocene", "Gelasian(LowPleio2)", "Upper Pliocene",
#                                     "Lower Pliocene 1", "Lower Pliocene 2","Upper Miocene 1",
#                                     "Upper Miocene 2", "Upper Miocene 3","Middle Miocene 1", 
#                                     "Middle Miocene 2","Lower Miocene 1", "Lower Miocene 2"))



HistCLBins <- PleiPlioCL %>%
 # merge(BINS) %>%
  filter( !is.na(CL)) %>%
  mutate(CL=log(CL)) %>%
  arrange(MeanBins) %>%
#  filter(EpochBins != "Modern") %>%
  ggplot(aes(CL)) + 
  geom_histogram( aes(y=..density..),col="black", fill="gray")+
  #geom_density() +
  geom_density(alpha=.2, fill="darkslategray1") +
  facet_wrap(~EpochBins, scales="free_x")+   #binwidth=10,
  theme_classic()+ 
  #theme_gray() +
  theme(panel.border = element_rect(colour = "black", fill=NA))


HistCLBins


StatsBins <- PleiPlioCL %>%
  filter(!is.na(CL)) %>%
  group_by(EpochBins) %>%
  summarise(nCL=length(CL), min=min(CL), max=max(CL), var=var(CL), mean=round(mean(CL),1), logm=round(mean(log10(CL)),1),
            med=round(median(CL),1), logmed=round(median(log10(CL)),1), #CLmode=mode(CL), CLlogmode=mode(log10(CL)),
            skew=round(skewness(CL),2), logsk=round(skewness(log10(CL)),2),
            kurt=round(kurtosis(CL),2), logku=round(kurtosis(log10(CL)),2) ) %>%
  unique()

#kable(StatsBins)


```





\newpage

## modern vs. fossil

```{r "Histograms of body size data, modern vs. fossil", echo=FALSE, fig.cap="Distribution of body size data modern vs. fossil, logtransformed."}


# Epoch <- as.vector(c("Modern", "Pleistocene", "Pleistocene", "Pleistocene", 
#                       "Pliocene", "Pliocene", "Miocene", "Miocene", "Miocene", "Oligocene and Eocene"))
# EPOCH <- as.vector(c("Modern", "Fossil", "Fossil", "Fossil"
#                      , "Fossil", "Fossil", "Fossil", "Fossil", "Fossil", "Fossil"))

Epoch <- as.vector(c("Modern", "Pleistocene", "Pleistocene", "Pleistocene", "Pleistocene", 
                      "Pliocene", "Pliocene","Pliocene",
                     "Miocene", "Miocene", "Miocene",  "Miocene", "Miocene", "Miocene", "Miocene"))
EPOCH <- as.vector(c("Modern", "Fossil", "Fossil", "Fossil"
                     , "Fossil", "Fossil", "Fossil", "Fossil", "Fossil", "Fossil"
                     , "Fossil", "Fossil", "Fossil", "Fossil", "Fossil"))
EpochBins <- as.vector(unique(PleiPlioCL$EpochBins))
Epochs <- data.frame(Epoch,EpochBins, EPOCH)


Epochs$EPOCH = factor(Epochs$EPOCH,levels=c("Modern", "Fossil"))

HistCLFossil <- PleiPlioCL %>%
  merge(Epochs) %>%
  filter( !is.na(CL)) %>%
  mutate(CL=log(CL)) %>%
#  filter(EpochBins != "Modern") %>%
  ggplot(aes(CL)) + 
  geom_histogram( aes(y=..density..),col="black", fill="gray")+
  #geom_density() +
  geom_density(alpha=.2, fill="darkslategray1") +
  facet_wrap(~EPOCH, scales="free_x")+   #binwidth=10,
  theme_classic()+ 
  #theme_gray() +
  theme(panel.border = element_rect(colour = "black", fill=NA))


HistCLFossil


StatsFossil <- PleiPlioCL %>%
  merge(Epochs) %>%
  filter(!is.na(CL)) %>%
  group_by(EPOCH) %>%
  summarise(nCL=length(CL), min=min(CL), max=max(CL), var=var(CL), mean=round(mean(CL),1), logm=round(mean(log10(CL)),1),
            med=round(median(CL),1), logmed=round(median(log10(CL)),1), #CLmode=mode(CL), CLlogmode=mode(log10(CL)),
            skew=round(skewness(CL),2), logsk=round(skewness(log10(CL)),2),
            kurt=round(kurtosis(CL),2), logku=round(kurtosis(log10(CL)),2) ) %>%
  unique()

#kable(StatsFossil)


```


\newpage

## modern vs. fossil, continental vs. insular


```{r "Histograms of body size data, modern vs. fossil, continental vs. insular", echo=FALSE, fig.cap="Distribution of body size data modern vs. fossil, continental vs. insular logtransformed."}


# Epoch <- as.vector(c("Modern", "Pleistocene", "Pleistocene", "Pleistocene", 
#                       "Pliocene", "Pliocene", "Miocene", "Miocene", "Miocene", "Oligocene and Eocene"))
# EPOCH <- as.vector(c("Modern", "Fossil", "Fossil", "Fossil"
#                      , "Fossil", "Fossil", "Fossil", "Fossil", "Fossil", "Fossil"))
# EpochBins <- as.vector(unique(PleiPlioCL$EpochBins))
# Epochs <- data.frame(Epoch,EpochBins, EPOCH)


HistCLFossilIsland <- PleiPlioCL %>%
  merge(Epochs) %>%
  filter( !is.na(CL)) %>%
  mutate(CL=log(CL)) %>%
#  filter(EpochBins != "Modern") %>%
  ggplot(aes(CL)) + 
  geom_histogram( aes(y=..density..),col="black", fill="gray")+
  #geom_density() +
  geom_density(alpha=.2, fill="darkslategray1") + 
  facet_wrap(EPOCH~Island, scales="free_x")+   #binwidth=10,
  theme_classic()+ 
  #theme_gray() +
  theme(panel.border = element_rect(colour = "black", fill=NA))


HistCLFossilIsland


StatsFossilIsland <- PleiPlioCL %>%
  merge(Epochs) %>%
  filter(!is.na(CL)) %>%
  group_by(EPOCH, Island) %>%
  summarise(nCL=length(CL), min=min(CL), max=max(CL), var=var(CL), mean=round(mean(CL),1), logm=round(mean(log10(CL)),1),
            med=round(median(CL),1), logmed=round(median(log10(CL)),1), #CLmode=mode(CL), CLlogmode=mode(log10(CL)),
            skew=round(skewness(CL),2), logsk=round(skewness(log10(CL)),2),
            kurt=round(kurtosis(CL),2), logku=round(kurtosis(log10(CL)),2) ) %>%
  unique()

#kable(StatsFossilIsland)


```

\newpage
## continental vs. insular

```{r "Histograms of body size data, continental vs. insular", echo=FALSE, fig.cap="Distribution of body site data of continental (n) and insular(y) species, logtransformed."}


HistCLIsland <- PleiPlioCL %>%
  filter( !is.na(CL)) %>%
  mutate(CL=log(CL)) %>%
#  filter(EpochBins != "Modern") %>%
  ggplot(aes(CL)) + 
  geom_histogram( aes(y=..density..),col="black", fill="gray")+
  #geom_density() +
  geom_density(alpha=.2, fill="darkslategray1") + 
  facet_wrap(~Island, scales="free_x")+   #binwidth=10,
  theme_classic()+ 
  #theme_gray() +
  theme(panel.border = element_rect(colour = "black", fill=NA))

#HistCLModern
HistCLIsland


StatsIsland <- PleiPlioCL %>%
  filter(!is.na(CL)) %>%
  group_by(Island) %>%
  summarise(nCL=length(CL), min=min(CL), max=max(CL), var=var(CL), mean=round(mean(CL),1), logm=round(mean(log10(CL)),1),
            med=round(median(CL),1), logmed=round(median(log10(CL)),1), #CLmode=mode(CL), CLlogmode=mode(log10(CL)),
            skew=round(skewness(CL),2), logsk=round(skewness(log10(CL)),2),
            kurt=round(kurtosis(CL),2), logku=round(kurtosis(log10(CL)),2) ) %>%
  unique()

#kable(StatsIsland)
```

\newpage
## continents

```{r "Histograms of body size data, split by continents", echo=FALSE, fig.cap="Distribution of body site data per continent, logtransformed."}

Continent <- as.vector(unique(PleiPlioCL$Continent))
Con <- as.vector(c("Asia", "Africa", "Europe", "America", "America", "America", "Europe"))#
Continents <- data.frame(Continent, Con)

HistCLContinents <- PleiPlioCL %>%
  filter( !is.na(CL)) %>%
  mutate(CL=log(CL)) %>%
  merge(Continents) %>%
#  filter(EpochBins != "Modern") %>%
  ggplot(aes(CL)) + 
  geom_histogram( aes(y=..density..),col="black", fill="gray")+
  #geom_density() +
  geom_density(alpha=.2, fill="darkslategray1") +
  facet_wrap(~Con, scales="free_x")+   #binwidth=10,
  theme_classic()+ 
  #theme_gray() +
  theme(panel.border = element_rect(colour = "black", fill=NA))


HistCLContinents


StatsContinents <- PleiPlioCL %>%
  merge(Continents) %>%
  filter(!is.na(CL)) %>%
  group_by(Con) %>%
  summarise(nCL=length(CL), min=min(CL), max=max(CL), var=var(CL), mean=round(mean(CL),1), logm=round(mean(log10(CL)),1),
            med=round(median(CL),1), logmed=round(median(log10(CL)),1), #CLmode=mode(CL), CLlogmode=mode(log10(CL)),
            skew=round(skewness(CL),2), logsk=round(skewness(log10(CL)),2),
            kurt=round(kurtosis(CL),2), logku=round(kurtosis(log10(CL)),2) ) %>%
  unique()

#kable(StatsContinents)
```


```{r "Histograms of body size data, genera", echo=FALSE, fig.cap="Distribution of body size, genera, logtransformed.", include=FALSE}

Continent <- as.vector(unique(PleiPlioCL$Continent))
Con <- as.vector(c("Asia", "Africa", "Europe", "America", "America", "America", "Europe"))#
Continents <- data.frame(Continent, Con)

HistCLGenera <- PleiPlioCL %>%
  filter( !is.na(CL)) %>%
  mutate(CL=log(CL)) %>%
  merge(Continents) %>%
#  filter(EpochBins != "Modern") %>%
  ggplot(aes(CL)) + 
  geom_histogram( aes(y=..density..),col="black", fill="gray")+
  #geom_density() +
  geom_density(alpha=.2, fill="darkslategray1") + 
  facet_wrap(~Genus, scales="free_x")+   #binwidth=10,
  theme_classic()+ 
  #theme_gray() +
  theme(panel.border = element_rect(colour = "black", fill=NA))


HistCLGenera


StatsGenera <- PleiPlioCL %>%
  merge(Continents) %>%
  filter(!is.na(CL)) %>%
  group_by(Genus) %>%
  summarise(nCL=length(CL), min=min(CL), max=max(CL), var=var(CL), mean=round(mean(CL),1), logm=round(mean(log10(CL)),1),
            med=round(median(CL),1), logmed=round(median(log10(CL)),1), #CLmode=mode(CL), CLlogmode=mode(log10(CL)),
            skew=round(skewness(CL),2), logsk=round(skewness(log10(CL)),2),
            kurt=round(kurtosis(CL),2), logku=round(kurtosis(log10(CL)),2) ) %>%
  unique()

kable(StatsGenera)
```


\newpage
## General statistics

```{r "Tabel Stats Distribution CL", echo=FALSE}
Stats <- bind_rows(StatsAll, StatsBins, StatsFossil, StatsIsland, 
                   StatsFossilIsland, StatsContinents) %>%
  select(-EpochBins, -Island, -Con, -EPOCH)


#names(Stats) <- c("nCL", "min", "max", "var", "mean", "logm", "med", "logmed", "skew", "logsk", "kurt", "logku", "Variable")
#Stats[1, 13] <- 0
#Stats[is.na(Stats)] <-0# c("all", "continental", "insular")

# Stats$Variable <- c("all", "Modern", "Upper Pleistocene", "Middle Pleistocene", "Lower Pleistocene", "Upper Pliocene", "Lower Pliocene", "Upper Miocene", "Middle Miocene", "Lower Miocene", "Oligocene and Eocene","Fossil", "Modern", "continental", "insular", "fossil-con", "fossil-ins", "modern-con", 
#                     "modern-ins", "Africa", "America", "Asia", "Europe")


Stats$Variable <- c("all", "Modern", "Upper Pleistocene", "Middle Pleistocene", "Lower Pleistocene",
                  "Gelasian(LowPlei2)",  "Upper Pliocene", "Lower Pliocene 1", "Lower Pliocene 2",
                  "Upper Miocene 1","Upper Miocene 2","Upper Miocene 3", "Middle Miocene 1",
                  "Middle Miocene 2","Lower Miocene 1", "Lower Miocene 2", "Modern","Fossil", "continental", "insular", "fossil-con", "fossil-ins", "modern-con",
                    "modern-ins", "Africa", "America", "Asia", "Europe")


kable(Stats, caption="General statistics of body size data: all, per time bin, insular and continental, per continent (all referring to CL: min, max, variance, mean, logmean, median, logmedian, skewness, logskewness, kurosis, logkurtosis")

```

\newpage
# Boxplots
## genera per time bins

```{r "Boxplots of each genus per time bin", echo=FALSE, warning=FALSE, fig.cap="Boxplots of mean CL per time bin, including mean and sd CL for each genus (as pointrange)."}
# PleiPlioCL$EpochBins = factor(PleiPlioCL$EpochBins, levels=c("Modern", "Upper Pleistocene",
#                                                              "Middle Pleistocene", "Lower Pleistocene",
#                                                              "Upper Pliocene", "Lower Pliocene",
#                                                              "Upper Miocene", "Middle Miocene",
#                                                              "Lower Miocene", "Oligocene and Eocene"))
# 
# 
# Epoch <- as.vector(c("Modern", "Pleistocene", "Pleistocene", "Pleistocene", 
#                       "Pliocene", "Pliocene", "Miocene", "Miocene", "Miocene", "Oligocene and Eocene"))
# EpochBins <- as.vector(unique(PleiPlioCL$EpochBins))
# Epochs <- data.frame(Epoch,EpochBins )

# Continent <- as.vector(unique(PleiPlioCL$Continent))
# Con <- as.vector(c("Asia", "Africa", "Europe", "America", "America", "America", "Europe")) #
# Continents <- data.frame(Continent, Con)





IndGenera <- PleiPlioCL %>%
  group_by(EpochBins, Genus) %>%
  filter(!is.na(CL)) %>%
  summarise(GenusMean=mean(CL), GenusSD=sd(CL), n=n()) 

IndGenera[is.na(IndGenera)] <-0


give.n <- function(x){
  return(c(y = 0, label = length(x)))
}



BinGenera <- PleiPlioCL %>%
  group_by(EpochBins, Genus) %>%
  filter(!is.na(CL)) %>%
  summarise(GenusMean=mean(CL), GenusSD=sd(CL), n=n()) %>%
#  summarise(TimeBinMean=mean(GenusMean), TimeBinSD=sd(GenusMean), n=n()) %>%
  ungroup() %>%#, CL=CL
  ggplot(aes(EpochBins, GenusMean)) + #, colour=Genus
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  stat_boxplot(geom ='errorbar', width = 0.2) +
  geom_boxplot() +
  geom_pointrange(data=IndGenera, position="jitter", aes(x=EpochBins, y=GenusMean, colour=EpochBins, ymin = GenusMean-GenusSD, ymax = GenusMean+GenusSD)) +
  stat_summary(fun.data = give.n, geom = "text") +
  theme(legend.background = element_rect(colour = 'black'),
        panel.border = element_rect(colour = "black", fill=NA)) +
  scale_colour_manual(values=c("#808080", "#848484", "#878787", "#8B8B8B", "#8F8F8F", "#929292",
                               "#969696", "#999999", "#969696", "#999999",
                               "#969696", "#999999", "#969696", "#999999", "#999999")) +
  theme(legend.position="none")

BinGenera

```

\newpage
## continental vs. insular per time bin

```{r "Boxplots of each genus per time bin, continental vs. insular", echo=FALSE, warning=FALSE, fig.cap="Boxplots of each genus per time bin, continental vs. insular species."}
# PleiPlioCL$EpochBins = factor(PleiPlioCL$EpochBins, levels=c("Modern", "Upper Pleistocene",
#                                                              "Middle Pleistocene", "Lower Pleistocene",
#                                                              "Upper Pliocene", "Lower Pliocene",
#                                                              "Upper Miocene", "Middle Miocene",
#                                                              "Lower Miocene", "Oligocene and Eocene"))
# 
# 
# Epoch <- as.vector(c("Modern", "Pleistocene", "Pleistocene", "Pleistocene", 
#                       "Pliocene", "Pliocene", "Miocene", "Miocene", "Miocene", "Oligocene and Eocene"))
# EpochBins <- as.vector(unique(PleiPlioCL$EpochBins))
# Epochs <- data.frame(Epoch,EpochBins )
# 
# Continent <- as.vector(unique(PleiPlioCL$Continent))
# Con <- as.vector(c("Asia", "Africa", "Europe", "America", "America", "America", "Europe"))#
# Continents <- data.frame(Continent, Con)





IndGenera <- PleiPlioCL %>%
  group_by(EpochBins, Genus, Island) %>%
  filter(!is.na(CL)) %>%
  summarise(GenusMean=mean(CL), GenusSD=sd(CL), n=n()) 

IndGenera[is.na(IndGenera)] <- 0


give.n <- function(x){
  return(c(y = 0, label = length(x)))
}



BinGenera <- PleiPlioCL %>%
  group_by(EpochBins, Genus, Island) %>%
  filter(!is.na(CL)) %>%
  summarise(GenusMean=mean(CL), GenusSD=sd(CL), n=n()) %>%
#  summarise(TimeBinMean=mean(GenusMean), TimeBinSD=sd(GenusMean), n=n()) %>%
  ungroup() %>%#, CL=CL
  ggplot(aes(EpochBins, GenusMean)) + #, colour=Genus
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  stat_boxplot(geom ='errorbar', width = 0.2) +
  geom_boxplot() +
  geom_pointrange(data=IndGenera, position="jitter", aes(x=EpochBins, y=GenusMean, colour=EpochBins, ymin = GenusMean-GenusSD, ymax = GenusMean+GenusSD)) +
  facet_wrap(~Island)+
  stat_summary(fun.data = give.n, geom = "text") +
  theme(legend.background = element_rect(colour = 'black'),
        panel.border = element_rect(colour = "black", fill=NA)) +
  scale_colour_manual(values=c("#808080", "#848484", "#878787", "#8B8B8B", "#8F8F8F", "#929292",
                               "#969696", "#999999", "#969696", "#999999",
                               "#969696", "#999999", "#969696", "#999999", "#999999")) +
  theme(legend.position="none")

BinGenera

```

\newpage
## fossil vs. modern

```{r "Boxplots modern vs. fossil", echo=FALSE, warning=FALSE, fig.cap="Boxplots fossil vs. modern."}
# PleiPlioCL$EpochBins = factor(PleiPlioCL$EpochBins, levels=c("Modern", "Upper Pleistocene",
#                                                              "Middle Pleistocene", "Lower Pleistocene",
#                                                              "Upper Pliocene", "Lower Pliocene",
#                                                              "Upper Miocene", "Middle Miocene",
#                                                              "Lower Miocene", "Oligocene and Eocene"))
# 
# 
# Epoch <- as.vector(c("Modern", "Pleistocene", "Pleistocene", "Pleistocene", 
#                       "Pliocene", "Pliocene", "Miocene", "Miocene", "Miocene", "Oligocene and Eocene"))
# EPOCH <- as.vector(c("Modern", "Fossil", "Fossil", "Fossil"
#                      , "Fossil", "Fossil", "Fossil", "Fossil", "Fossil", "Fossil"))
# EpochBins <- as.vector(unique(PleiPlioCL$EpochBins))
# Epochs <- data.frame(Epoch,EpochBins, EPOCH)
# 
# Continent <- as.vector(unique(PleiPlioCL$Continent))
# Con <- as.vector(c("Asia", "Africa", "Europe", "America", "America", "America", "Europe"))#
# Continents <- data.frame(Continent, Con)





IndGenera <- PleiPlioCL %>%
  merge(Epochs) %>%
  group_by(EPOCH, Genus, Island) %>%
  filter(!is.na(CL)) %>%
  summarise(GenusMean=mean(CL), GenusSD=sd(CL), n=n()) 

IndGenera[is.na(IndGenera)] <- 0


give.n <- function(x){
  return(c(y = 0, label = length(x)))
}



BinGenera <- PleiPlioCL %>%
  merge(Epochs) %>%
  group_by(EPOCH, Genus, Island) %>%
  filter(!is.na(CL)) %>%
  summarise(GenusMean=mean(CL), GenusSD=sd(CL), n=n()) %>%
#  summarise(TimeBinMean=mean(GenusMean), TimeBinSD=sd(GenusMean), n=n()) %>%
  ungroup() %>%#, CL=CL
  ggplot(aes(EPOCH, GenusMean)) + #, colour=Genus
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  stat_boxplot(geom ='errorbar', width = 0.2) +
  geom_boxplot() +
  geom_pointrange(data=IndGenera, position="jitter", aes(x=EPOCH, y=GenusMean, colour=EPOCH, ymin = GenusMean-GenusSD, ymax = GenusMean+GenusSD)) +
#  facet_wrap(~Island)+
  stat_summary(fun.data = give.n, geom = "text") +
  theme(legend.background = element_rect(colour = 'black'),
        panel.border = element_rect(colour = "black", fill=NA)) +
  scale_colour_manual(values=c("#808080", "#848484", "#878787", "#8B8B8B", "#8F8F8F", "#929292",
                               "#969696", "#999999", "#969696", "#999999",
                               "#969696", "#999999", "#969696", "#999999", "#999999")) +
  theme(legend.position="none")

BinGenera




Modern <- PleiPlioCL$CL[which(PleiPlioCL$EpochBins=="Modern")]
Fossil <- PleiPlioCL$CL[which(PleiPlioCL$EpochBins!="Modern")]



ModernFossil<- wilcox.test(Modern,Fossil, alternative = "l",paired=FALSE)  

#ModernFossil

#ModernFossil$p.value

#ModernFossil$p.value < 0.05



```

Wilcoxon Rank Sum Test (unpaired data): 

modern < fossil (P = $`r ModernFossil$p.value`$)


\newpage
## fossil vs. modern, continental vs. insular


```{r "Boxplots fossil vs. modern, continental vs. insular", echo=FALSE, warning=FALSE, fig.cap="Boxplots fossil vs. modern, continental vs. insular species."}
# PleiPlioCL$EpochBins = factor(PleiPlioCL$EpochBins, levels=c("Modern", "Upper Pleistocene",
#                                                              "Middle Pleistocene", "Lower Pleistocene",
#                                                              "Upper Pliocene", "Lower Pliocene",
#                                                              "Upper Miocene", "Middle Miocene",
#                                                              "Lower Miocene", "Oligocene and Eocene"))
# 
# 
# Epoch <- as.vector(c("Modern", "Pleistocene", "Pleistocene", "Pleistocene", 
#                       "Pliocene", "Pliocene", "Miocene", "Miocene", "Miocene", "Oligocene and Eocene"))
# EPOCH <- as.vector(c("Modern", "Fossil", "Fossil", "Fossil"
#                      , "Fossil", "Fossil", "Fossil", "Fossil", "Fossil", "Fossil"))
# EpochBins <- as.vector(unique(PleiPlioCL$EpochBins))
# Epochs <- data.frame(Epoch,EpochBins, EPOCH)
# 
# Continent <- as.vector(unique(PleiPlioCL$Continent))
# Con <- as.vector(c("Asia", "Africa", "Europe", "America", "America", "America", "Europe"))#
# Continents <- data.frame(Continent, Con)





IndGenera <- PleiPlioCL %>%
  merge(Epochs) %>%
  group_by(EPOCH, Genus, Island) %>%
  filter(!is.na(CL)) %>%
  summarise(GenusMean=mean(CL), GenusSD=sd(CL), n=n()) 

IndGenera[is.na(IndGenera)] <- 0


give.n <- function(x){
  return(c(y = 0, label = length(x)))
}



BinGenera <- PleiPlioCL %>%
  merge(Epochs) %>%
  group_by(EPOCH, Genus, Island) %>%
  filter(!is.na(CL)) %>%
  summarise(GenusMean=mean(CL), GenusSD=sd(CL), n=n()) %>%
#  summarise(TimeBinMean=mean(GenusMean), TimeBinSD=sd(GenusMean), n=n()) %>%
  ungroup() %>%#, CL=CL
  ggplot(aes(EPOCH, GenusMean)) + #, colour=Genus
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  stat_boxplot(geom ='errorbar', width = 0.2) +
  geom_boxplot() +
  geom_pointrange(data=IndGenera, position="jitter", aes(x=EPOCH, y=GenusMean, colour=EPOCH, ymin = GenusMean-GenusSD, ymax = GenusMean+GenusSD)) +
  facet_wrap(~Island)+
  stat_summary(fun.data = give.n, geom = "text") +
  theme(legend.background = element_rect(colour = 'black'),
        panel.border = element_rect(colour = "black", fill=NA)) +
  scale_colour_manual(values=c("#808080", "#848484", "#878787", "#8B8B8B", "#8F8F8F", "#929292",
                               "#969696", "#999999", "#969696", "#999999",
                               "#969696", "#999999", "#969696", "#999999", "#999999")) +
  theme(legend.position="none")

BinGenera




ModernIsland <- PleiPlioCL$CL[which(PleiPlioCL$EpochBins=="Modern" & PleiPlioCL$Island=="y")]
FossilIsland <- PleiPlioCL$CL[which(PleiPlioCL$EpochBins!="Modern" & PleiPlioCL$Island=="y")]


ModernFossilIsland<- wilcox.test(ModernIsland,FossilIsland, alternative = "l",paired=FALSE)  

# ModernFossilIsland
# 
# ModernFossilIsland$p.value
# 
# ModernFossilIsland$p.value < 0.05



ModernCon <- PleiPlioCL$CL[which(PleiPlioCL$EpochBins=="Modern" & PleiPlioCL$Island=="n")]
FossilCon <- PleiPlioCL$CL[which(PleiPlioCL$EpochBins!="Modern" & PleiPlioCL$Island=="n")]


ModernFossilCon<- wilcox.test(ModernCon,FossilCon, alternative = "l",paired=FALSE)  

# ModernFossilCon
# 
# ModernFossilCon$p.value
# 
# ModernFossilCon$p.value < 0.05


```
Wilcoxon Rank Sum Test (unpaired data): 

modern continental < fossil continental (P = $`r ModernFossilCon$p.value`$)

modern insular < fossil insular (P = $`r ModernFossilIsland$p.value`$)




```{r "Boxplot continental vs. insular, individuals", echo=FALSE, warning=FALSE, fig.cap="Boxplot continental vs. insular, individuals", include=FALSE}

# Epoch <- as.vector(c("Modern", "Pleistocene", "Pleistocene", "Pleistocene",
#                       "Pliocene", "Pliocene", "Miocene", "Miocene", "Miocene", "Oligocene and Eocene"))
# EpochBins <- as.vector(unique(PleiPlioCL$EpochBins))
# Epochs <- data.frame(Epoch,EpochBins )
# 
# 
# Epochs$EpochBins = factor(Epochs$EpochBins, levels=c("Modern", "Pleistocene",
#                                                              "Pliocene",
#                                                              "Miocene", "Oligocene and Eocene"))
# 
# 
# Continent <- as.vector(unique(PleiPlioCL$Continent))
# Con <- as.vector(c("Asia", "Africa", "Europe", "America", "America", "America", "Europe"))#
# Continents <- data.frame(Continent, Con)

Island <- PleiPlioCL %>%
  dplyr::select(Taxon, CL, PL, Age, Island, Continent, EpochBins) %>%
  merge(Epochs) %>%
  merge(Continents) %>%
  dplyr::select(CL, Island, Con, Epoch, Age)
  

Island$Epoch = factor(Island$Epoch, levels=c("Modern", "Pleistocene",
                                                             "Pliocene",
                                                             "Miocene", "Oligocene and Eocene"))

colnames(Island)[3] <- "Continent"

give.n <- function(x){
  return(c(y = 0, label = length(x)))
}


IslandBox <- Island %>%
  dplyr::select(CL, Island, Continent, Epoch, Age) %>%
  ggplot(aes(Island, CL)) + geom_boxplot() + geom_jitter(aes(colour=Continent, shape=Epoch)) +
#  facet_grid(.~Continent) +
  theme_classic() + # for white background 
#  stat_summary(fun.data = give.n, geom = "text") +
 # xlab("Sex") + ylab("Carapace length [mm]") + 
  theme(legend.background = element_rect(colour = 'black'),
        panel.border = element_rect(colour = "black", fill=NA)) +
  stat_summary(fun.data = give.n, geom = "text") +
  scale_colour_brewer(palette="Set1")

IslandBox




```



```{r "Boxplots, species summarised", echo=FALSE, fig.cap="Boxplots continental vs. insular, species summarised", include=FALSE}
Island <- PleiPlioCL %>%
  dplyr::select(Taxon, CL, PL, Age, Island, Continent, EpochBins) %>%
  merge(Epochs) %>%
  merge(Continents) %>%
  dplyr::select(CL, Island, Con, Epoch, Age, Taxon)
  
colnames(Island)[3] <- "Continent"

give.n <- function(x){
  return(c(y = 0, label = length(x)))
}


IslandBoxSpecies <- Island %>%
  dplyr::select(CL, Island, Continent, Epoch, Age, Taxon) %>%
  group_by(Taxon, Island, Continent, Epoch) %>%
  dplyr::filter(!is.na(CL)) %>%
  summarise(meanCL=mean(CL)) %>%
  ggplot(aes(Island, meanCL)) + geom_boxplot() + geom_jitter(aes(colour=Continent, shape=Epoch)) +
#  facet_grid(.~Continent) +
  theme_classic() + # for white background 
#  stat_summary(fun.data = give.n, geom = "text") +
 # xlab("Sex") + ylab("Carapace length [mm]") + 
  theme(legend.background = element_rect(colour = 'black'),
        panel.border = element_rect(colour = "black", fill=NA)) +
  stat_summary(fun.data = give.n, geom = "text") +
  scale_colour_brewer(palette="Set1")

IslandBoxSpecies
```


\newpage
## continental vs. insular

```{r "Boxplot continental vs. insular", echo=FALSE, fig.cap="Boxplot continental vs. insular, genera summarised"}


give.n <- function(x){
  return(c(y = 0, label = length(x)))
}


IslandGenus <- testRatio %>%
  filter(!is.na(CL)) %>%
  select(CL, Island, Genus) %>%
  filter(!is.na(CL)) %>%
  group_by(Genus, Island) %>%
  summarise(meanCL=mean(CL), sdCL=sd(CL))

IslandGenus[is.na(IslandGenus)] <- 0


IslandAll <- testRatio %>%
  select(CL, Island, Genus) %>%
  filter(!is.na(CL)) %>%
  ggplot(aes(Island, CL)) + geom_boxplot() + 
  geom_pointrange(data=IslandGenus, position="jitter", aes(x=Island, y=meanCL,  colour= Island , ymin = meanCL-sdCL, ymax = meanCL+sdCL)) +   #colour=Genus,
  theme_classic() + # for white background 
  stat_summary(fun.data = give.n, geom = "text") +
  theme(legend.background = element_rect(colour = 'black'),
        panel.border = element_rect(colour = "black", fill=NA)) +
  scale_colour_manual(values=c("#808080", "#848484", "#878787", "#8B8B8B", "#8F8F8F", "#929292",
                               "#969696", "#999999")) +
  theme(legend.position="none")

IslandAll





Insular <- PleiPlioCL$CL[which(PleiPlioCL$Island=="y")]
Continental <- PleiPlioCL$CL[which(PleiPlioCL$Island=="n")]



InsularContinental<- wilcox.test(Insular,Continental, alternative = "g",paired=FALSE)  

# InsularContinental
# 
# InsularContinental$p.value
# 
# InsularContinental$p.value < 0.05



```

Wilcoxon Rank Sum Test (unpaired data): 

continental < insular (P = $`r InsularContinental$p.value`$)

\newpage
## continental vs. insular per time bin

```{r "Boxplot continental vs. insular, split into time bins", echo=FALSE, fig.cap="Boxplot continental vs. insular, genera summarised"}


give.n <- function(x){
  return(c(y = 0, label = length(x)))
}


IslandGenus <- PleiPlioCL %>%
  select(CL, Island, Genus, EpochBins) %>%
  filter(!is.na(CL)) %>%
  group_by(Genus, Island, EpochBins) %>%
  summarise(meanCL=mean(CL), sdCL=sd(CL))

IslandGenus[is.na(IslandGenus)] <- 0


IslandAll <- PleiPlioCL %>%
  select(CL, Island, Genus, EpochBins) %>%
  filter(!is.na(CL)) %>%
  ggplot(aes(Island, CL)) + geom_boxplot() + 
  geom_pointrange(data=IslandGenus, position="jitter", aes(x=Island, y=meanCL,  colour= Island , ymin = meanCL-sdCL, ymax = meanCL+sdCL)) +   #colour=Genus,
  theme_classic() + # for white background 
  stat_summary(fun.data = give.n, geom = "text") +
  theme(legend.background = element_rect(colour = 'black'),
        panel.border = element_rect(colour = "black", fill=NA)) +
  facet_wrap(~EpochBins) +
  scale_colour_manual(values=c("#808080", "#848484", "#878787", "#8B8B8B", "#8F8F8F", "#929292",
                               "#969696", "#999999")) +
  theme(legend.position="none")

IslandAll

```

\newpage
## continents

```{r "Boxplot body size split into continents", echo=FALSE, fig.cap="Boxplot: body size on different continents, genera summarised"}


give.n <- function(x){
  return(c(y = 0, label = length(x)))
}


ContinentGenus <- testRatio %>%
  merge(Continents) %>%
  select(CL, Con, Genus) %>%
  filter(!is.na(CL)) %>%
  group_by(Genus, Con) %>%
  summarise(meanCL=mean(CL), sdCL=sd(CL))


ContinentGenus[is.na(ContinentGenus)] <- 0


ContinentAll <- PleiPlioCL %>%
  merge(Continents) %>%
  select(CL, Con, Genus) %>%
  filter(!is.na(CL)) %>%
  ggplot(aes(Con, CL)) + geom_boxplot() + 
  geom_pointrange(data=ContinentGenus, position="jitter", aes(x=Con, y=meanCL,  colour= Con , ymin = meanCL-sdCL, ymax = meanCL+sdCL)) +   #colour=Genus,
  theme_classic() + # for white background 
  stat_summary(fun.data = give.n, geom = "text") +
  theme(legend.background = element_rect(colour = 'black'),
        panel.border = element_rect(colour = "black", fill=NA)) +
  scale_colour_manual(values=c("#808080", "#848484", "#878787", "#8B8B8B", "#8F8F8F", "#929292",
                               "#969696", "#999999")) +
  theme(legend.position="none")

ContinentAll

continent <- PleiPlioCL %>%
  merge(Continents)

Africa <- continent$CL[which(continent$Con=="Africa")]
America <- continent$CL[which(continent$Con=="America")]
Asia <- continent$CL[which(continent$Con=="Asia")]
Europe <- continent$CL[which(continent$Con=="Europe")]
Eurasia <- continent$CL[which(continent$Con=="Europe" | continent$Con=="Asia")]

ContinentsCompared <- kruskal.test(list(Africa, America, Eurasia))  #Asia, Europe

# ContinentsCompared
# 
# ContinentsCompared$p.value
# 
# ContinentsCompared$p.value < 0.05


```


Wilcoxon Rank Sum Test (unpaired data): 

Continent means differ (P = $`r ContinentsCompared$p.value`$)
(still have to look into the details...)



\newpage
## continents, continental vs. insular

```{r "Boxplot body size split into continents, continental vs. insular", echo=FALSE, fig.cap="Boxplot: body size on different continents, genera summarised"}


give.n <- function(x){
  return(c(y = 0, label = length(x)))
}


ContinentsGenus <- testRatio %>%
  merge(Continents) %>%
  select(CL, Con, Genus, Island) %>%
  filter(!is.na(CL)) %>%
  group_by(Genus, Con, Island) %>%
  summarise(meanCL=mean(CL), sdCL=sd(CL))


ContinentsGenus[is.na(ContinentsGenus)] <- 0


ContinentsAll <- testRatio %>%
  merge(Continents) %>%
  select(CL, Con, Genus, Island) %>%
  filter(!is.na(CL)) %>%
  ggplot(aes(Island, CL)) + geom_boxplot() + 
  geom_pointrange(data=ContinentsGenus, position="jitter", aes(x=Island, y=meanCL,  colour= Con , ymin = meanCL-sdCL, ymax = meanCL+sdCL)) +   #colour=Genus,
  theme_classic() + # for white background 
  facet_wrap(~Con) +
  stat_summary(fun.data = give.n, geom = "text") +
  theme(legend.background = element_rect(colour = 'black'),
        panel.border = element_rect(colour = "black", fill=NA)) +
  scale_colour_manual(values=c("#808080", "#848484", "#878787", "#8B8B8B", "#8F8F8F", "#929292",
                               "#969696", "#999999")) +
  theme(legend.position="none")

ContinentsAll

```



```{r "Boxplots continents, insular vs. continent with age indicated",  echo=FALSE, warning=FALSE, fig.cap="Boxplots of body size (individuals) on different continents, insular vs. continent with age indicated", include=FALSE}

Island <- PleiPlioCL %>%
  dplyr::select(Genus, CL, PL, Age, Island, Continent, EpochBins) %>%
  merge(Epochs) %>%
  merge(Continents) %>%
  dplyr::select(CL, Island, Continent, Epoch, Age, Genus) %>%
  filter(!is.na(CL)) #%>%
  
  
  
IslandJitter <- Island %>%
  group_by(Genus, Island, Continent) %>%
  summarise(GenusMean=mean(CL), GenusSD=sd(CL), n=n(), Age=mean(Age)) 

#colnames(Island)[3] <- "Continent"
Island[is.na(Island)]<-0




IslandBoxAge <- Island %>%
  ggplot(aes(Island, CL)) + geom_boxplot() +
#  geom_jitter(data=IslandJitter,aes(colour=Age)) +
  facet_grid(.~Continent) +
  theme_classic() + # for white background 
#  stat_summary(fun.data = give.n, geom = "text") +
 # xlab("Sex") + ylab("Carapace length [mm]") + 
  theme(legend.background = element_rect(colour = 'black'),
        panel.border = element_rect(colour = "black", fill=NA))# +
#  scale_colour_gradientn(colors=c("orange", "red", "purple", "blue", "green")) #

IslandBoxAge
```


```{r "Separate genera per time bin", echo=FALSE, warning=FALSE,  fig.cap="Mean body size and standard deviation per genus in each time bin", include=FALSE}
BinGeneraFacet <- PleiPlioCL %>%
  group_by(EpochBins, Genus) %>%
  filter(!is.na(CL)) %>%
  summarise(GenusMean=mean(CL), GenusSD=sd(CL), n=n()) %>%
#  summarise(TimeBinMean=mean(GenusMean), TimeBinSD=sd(GenusMean), n=n()) %>%
  ungroup() %>%#, CL=CL
  ggplot(aes(EpochBins, GenusMean)) + #, colour=Genus
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
#  stat_boxplot(geom ='errorbar', width = 0.2) +
#  geom_boxplot() +
  geom_point(data=IndGenera,  aes(x=Genus, y=GenusMean, colour=Genus)) +
  geom_pointrange(data=IndGenera, aes(x=Genus, y=GenusMean, colour=Genus, ymin = GenusMean-GenusSD, ymax = GenusMean+GenusSD)) +
#  geom_boxplot(data=IndGenera,  aes(x=Genus, y=GenusMean, colour=Genus))+
#  geom_text(data=IndGenera,aes(label=Genus, x=EpochBins, y=GenusMean)) +
  facet_wrap(~EpochBins, scales="free_x") +
  guides(colour="none")#+
#  stat_boxplot(data=BinGenera, geom ='errorbar', width = 0.2) +
#  geom_boxplot(data=BinGenera) 
#  geom_errorbar(data=IndGenera, position="jitter",aes(ymin = GenusMean-GenusSD, ymax = GenusMean+GenusSD, width=0.2, colour=Genus))#+
  #scale_color_brewer(palette="Set1")
  #scale_colour_gradientn(colours=rainbow(4))

BinGeneraFacet
```


\newpage

# paleoTS analysis

## all (continental and insular)

### genera (all)

```{r "paleoTS plot with genus mean, including island species", echo=FALSE, fig.cap="paleoTS plot with genus mean, including island species"}
GenusMean <- PleiPlioCL %>%
  filter(EpochBins != "Modern") %>%
  filter(!is.na(CL)) %>%
  group_by(EpochBins, Genus) %>%
  summarise(meanCL = mean(CL), n=n()) %>%
#  ungroup() %>%
#  group_by(EpochBins) %>%
  summarise(mm = mean(meanCL), nn=n(), vv=var(meanCL)) %>%
  merge(BINSMio) %>%
  select(mm, nn, vv, tt=MeanBins)
  
#na <- PleiPlioCL[!complete.cases(PleiPlioCL),]



GenusModern <- PleiPlioCL %>%
  filter(EpochBins == "Modern") %>%
  filter(!is.na(CL)) %>%
  group_by(Genus) %>%
  summarise(meanCL=mean(CL), sdCL=sd(CL), n=n(), Age=mean(Age), EpochBins=unique(EpochBins), MeanBins=unique(MeanBins)) #%>%
#  summarise(mm=mean(meanCL), nn=n(), vv=var(meanCL), tt=mean(MeanBins))



sumTort <- read.csv("tortoises_summary.csv", sep=";", header=TRUE)
#colnames(sumTort)[1] <- "Taxon"
colnames(sumTort)[7] <- "meanCL"
colnames(sumTort)[8] <- "sdCL"
#sumTort$EpochBins <- "Modern"
#sumTort$MeanBins <- 0.0000005


sumTortoises <- sumTort %>%
  mutate(Genus= as.character(Genus)) %>%
  mutate(MeanBins=(Mamin+Mamax)/2) %>%
  select(Genus, MeanBins,  meanCL, sdCL, n) %>% #n,
  bind_rows(GenusModern)


SumTort <- sumTortoises %>%
  group_by(Genus) %>%
  mutate(tt=MeanBins, vv=sdCL^2, nn=n, mm=meanCL) %>%
  dplyr::select(mm, nn, vv, tt, Genus) %>%
  mutate(nx = nn*mm) %>%
  mutate(mmall=sum(nx)/sum(nn)) %>%
  mutate(SD=sqrt(nx), d=mm-mmall) %>%
  mutate(nsd=((nx^2+d^2)*nn)) %>%
  mutate(varall=sum(nsd)/sum(nn), n=sum(nn)) %>%
  dplyr::select(mm=mmall, vv=varall, nn=n, tt, Genus) %>%
  unique() %>%
  dplyr::select(CL=mm, n=nn, var=vv, tt, Genus)

#write.table(SumTort,file="SumTortModernGenus.txt", sep="\t", row.names = FALSE)

#kable(SumTort, caption="Overview over body size means per time bin on genus level.")

#boxplot(SumTort$CL, caption="Body size distribution in time bin 'modern'.")


modernMeanGenus <- SumTort %>%
  ungroup() %>%
  select(CL, n, tt) %>%
  filter( !is.na(CL)) %>%
  group_by(tt) %>%
  summarise(mm = mean(CL), vv=var(CL), nn=n())

#bis hier alle modernen taxa zusammengefasst (summarised and MFN_testudinidae)


GenusPaleo <- modernMeanGenus %>%
  bind_rows(GenusMean)%>%
  arrange(tt)


#kable(GenusPaleo)


paleoGen <-as.paleoTS(GenusPaleo$mm, GenusPaleo$vv, GenusPaleo$nn, GenusPaleo$tt, MM = NULL, genpars = NULL, label = "Testudinidae body size evolution mode, Holocene-Upper Miocene, genus mean CL", reset.time=TRUE)

plot(paleoGen)

```



```{r "Model-fitting, genera" , echo=FALSE, message=FALSE, include=FALSE}

PaleoGenFit <- (fit3models(paleoGen, silent=FALSE, method="AD", pool=FALSE))

PaleoGenFit2 <- (fit4models(paleoGen, silent=FALSE, method="AD", pool=FALSE))
```



```{r "Model-fitting, genera, table", echo=FALSE}

kable(PaleoGenFit, caption="Model-fitting results for testudinidae, genera, including island species")

kable(PaleoGenFit2, caption="Model-fitting results for testudinidae (4 models), genera, including island species")


```




\newpage

## continental (excluding insular species)

### genera (continental)

```{r "paleoTS plot with genus mean, excluding island species", echo=FALSE, message=FALSE, fig.cap="paleoTS plot with genus mean, excluding island species"}
GenusMeanIs <- PleiPlioCL %>%
  filter(EpochBins != "Modern" & Island == "n") %>%
  filter(!is.na(CL)) %>%
  group_by(EpochBins, Genus) %>%
  summarise(meanCL = mean(CL), n=n()) %>%
#  ungroup() %>%
#  group_by(EpochBins) %>%
  summarise(mm = mean(meanCL), nn=n(), vv=var(meanCL)) %>%
  merge(BINSMio) %>%
  select(mm, nn, vv, tt=MeanBins)
  
#na <- PleiPlioCL[!complete.cases(PleiPlioCL),]



GenusModernIs <- PleiPlioCL %>%
  filter(EpochBins == "Modern" & Island == "n") %>%
  filter(!is.na(CL)) %>%
  group_by(Genus) %>%
  summarise(meanCL=mean(CL), sdCL=sd(CL), n=n(), Age=mean(Age), EpochBins=unique(EpochBins), MeanBins=unique(MeanBins)) #%>%
#  summarise(mm=mean(meanCL), nn=n(), vv=var(meanCL), tt=mean(MeanBins))



sumTort <- read.csv("tortoises_summary.csv", sep=";", header=TRUE)
colnames(sumTort)[7] <- "meanCL"
colnames(sumTort)[8] <- "sdCL"


sumTortoises <- sumTort %>%
  dplyr::filter(Island == "n") %>%
  mutate(Genus= as.character(Genus)) %>%
  mutate(MeanBins=(Mamin+Mamax)/2) %>%
  select(Genus, MeanBins,  meanCL, sdCL, n) %>% #n,
  bind_rows(GenusModernIs)


SumTort <- sumTortoises %>%
  group_by(Genus) %>%
  mutate(tt=MeanBins, vv=sdCL^2, nn=n, mm=meanCL) %>%
  dplyr::select(mm, nn, vv, tt, Genus) %>%
  mutate(nx = nn*mm) %>%
  mutate(mmall=sum(nx)/sum(nn)) %>%
  mutate(SD=sqrt(nx), d=mm-mmall) %>%
  mutate(nsd=((nx^2+d^2)*nn)) %>%
  mutate(varall=sum(nsd)/sum(nn), n=sum(nn)) %>%
  dplyr::select(mm=mmall, vv=varall, nn=n, tt, Genus) %>%
  unique() %>%
  dplyr::select(CL=mm, n=nn, var=vv, tt, Genus)

#write.table(SumTort,file="SumTortModernGenus.txt", sep="\t", row.names = FALSE)

#kable(SumTort, caption="Overview over body size means per time bin on genus level, excluding island species.")

#boxplot(SumTort$CL, caption="Body size distribution in time bin 'modern'.")


modernMeanGenusIs <- SumTort %>%
  ungroup() %>%
  select(CL, n, tt) %>%
  filter( !is.na(CL)) %>%
  group_by(tt) %>%
  summarise(mm = mean(CL), vv=var(CL), nn=n())

#bis hier alle modernen taxa zusammengefasst (summarised and MFN_testudinidae)


GenusPaleoIs <- modernMeanGenusIs %>%
  bind_rows(GenusMeanIs)%>%
  arrange(tt)

GenusPaleoIs[is.na(GenusPaleoIs)] <- 0

#kable(GenusPaleoIs)


paleoGenIs <-as.paleoTS(GenusPaleoIs$mm, GenusPaleoIs$vv, GenusPaleoIs$nn, GenusPaleoIs$tt, MM = NULL, genpars = NULL, label = "Testudinidae body size evolution mode, Holocene-Upper Miocene, genus mean CL", reset.time=TRUE)

plot(paleoGenIs)
```



```{r "Model-fitting, genera, continental" , echo=FALSE, message=FALSE, include=FALSE}

PaleoGenIsFit <- (fit3models(paleoGenIs, silent=FALSE, method="AD", pool=FALSE))


PaleoGenIsFit2 <- (fit4models(paleoGenIs, silent=FALSE, method="AD", pool=FALSE))

```



```{r "Model-fitting, genera, continental, table", echo=FALSE}

kable(PaleoGenIsFit, caption="Model-fitting results for testudinidae, genera, excluding insular species")


kable(PaleoGenIsFit2, caption="Model-fitting results for testudinidae, genera, excluding insular species")


```



\newpage


## insular (excluding continental)

### genera (insular)

```{r "paleoTS plot with genus mean, excluding continental species", echo=FALSE, message=FALSE, fig.cap="paleoTS plot with genus mean, only insular species"}

GenusMeanCon <- PleiPlioCL %>%
  filter(EpochBins != "Modern" & Island == "y") %>%  # & EpochBins != "Upper Miocene"
  filter(!is.na(CL)) %>%
  group_by(EpochBins, Genus) %>%
  summarise(meanCL = mean(CL), n=n()) %>%
#  ungroup() %>%
#  group_by(EpochBins) %>%
  summarise(mm = mean(meanCL), nn=n(), vv=var(meanCL)) %>%
  merge(BINSMio) %>%
  select(mm, nn, vv, tt=MeanBins)
  
GenusMeanCon[is.na(GenusMeanCon)]<-0



GenusModernCon <- PleiPlioCL %>%
  filter(EpochBins == "Modern" & Island == "y") %>%
  filter(!is.na(CL)) %>%
  group_by(Genus) %>%
  summarise(meanCL=mean(CL), sdCL=sd(CL), n=n(), Age=mean(Age), EpochBins=unique(EpochBins), MeanBins=unique(MeanBins)) #%>%
#  summarise(mm=mean(meanCL), nn=n(), vv=var(meanCL), tt=mean(MeanBins))

GenusModernCon[is.na(GenusModernCon)]<-0


sumTort <- read.csv("tortoises_summary.csv", sep=";", header=TRUE)
colnames(sumTort)[7] <- "meanCL"
colnames(sumTort)[8] <- "sdCL"


sumTortoisesCon <- sumTort %>%
  dplyr::filter(Island == "y") %>%
  mutate(Genus= as.character(Genus)) %>%
  mutate(MeanBins=(Mamin+Mamax)/2) %>%
  select(Genus, MeanBins,  meanCL, sdCL, n) %>% #n,
  bind_rows(GenusModernCon)


SumTortCon <- sumTortoisesCon %>%
  group_by(Genus) %>%
  mutate(tt=MeanBins, vv=sdCL^2, nn=n, mm=meanCL) %>%
  dplyr::select(mm, nn, vv, tt, Genus) %>%
  mutate(nx = nn*mm) %>%
  mutate(mmall=sum(nx)/sum(nn)) %>%
  mutate(SD=sqrt(nx), d=mm-mmall) %>%
  mutate(nsd=((nx^2+d^2)*nn)) %>%
  mutate(varall=sum(nsd)/sum(nn), n=sum(nn)) %>%
  dplyr::select(mm=mmall, vv=varall, nn=n, tt, Genus) %>%
  unique() %>%
  dplyr::select(CL=mm, n=nn, var=vv, tt, Genus)

#write.table(SumTort,file="SumTortModernGenus.txt", sep="\t", row.names = FALSE)

#kable(SumTort, caption="Overview over body size means per time bin on genus level, excluding island species.")

#boxplot(SumTort$CL, caption="Body size distribution in time bin 'modern'.")


modernMeanGenusCon <- SumTortCon %>%
  ungroup() %>%
  select(CL, n, tt) %>%
  filter( !is.na(CL)) %>%
  group_by(tt) %>%
  summarise(mm = mean(CL), vv=var(CL), nn=n())

#bis hier alle modernen taxa zusammengefasst (summarised and MFN_testudinidae)


GenusPaleoCon <- modernMeanGenusCon %>%
  bind_rows(GenusMeanCon)%>%
  arrange(tt)


GenusPaleoCon[is.na(GenusPaleoCon)]<-0

#kable(GenusPaleoCon)


paleoGenCon <-as.paleoTS(GenusPaleoCon$mm, GenusPaleoCon$vv, GenusPaleoCon$nn, GenusPaleoCon$tt, MM = NULL, genpars = NULL, label = "Testudinidae body size evolution mode, Holocene-Upper Miocene, genus mean CL, insular genera", reset.time=TRUE)

plot(paleoGenCon)
```



```{r "Model-fitting, genera, insular" , echo=FALSE, message=FALSE, include=FALSE}

PaleoGenConFit <- (fit3models(paleoGenCon, silent=FALSE, method="AD", pool=FALSE))
```



```{r "Model-fitting, genera, insular, table", echo=FALSE}

kable(PaleoGenConFit, caption="Model-fitting results for testudinidae, genera, only insular species")

```




\newpage

## Equal time bins

### genera (equal bins)

```{r "Play around with time bins, generic level", echo=FALSE, fig.cap="Equal bins, genera"}

Alternative <- testRatio

Alternative$bin <- cut(Alternative$Age, c(0:50.000, breaks=0.0005))


AltBins <- data.frame(c( (0+0.0005)/2,(0.0005+1)/2,   (1+2)/2,  (2+3)/2, (3+4)/2,  (4+5)/2, (5+6)/2,   
(6+7)/2,  (7+8)/2,  (8+9)/2,  (9+10)/2,   
(10+11)/2,  (11+12)/2, (12+13)/2,  (13+14)/2,  (14+15)/2,  (15+16)/2,  (16+17)/2,   (17+18)/2,
(18+19)/2,  (19+20)/2,  (23+24)/2,   (25+26)/2,  (32+33)/2,   (33+34)/2,    (49+50)/2      ))

names(AltBins) <- "bins"



ABins <- Alternative %>%
  select(bin) %>%
  unique() %>%
  arrange(bin)


AltMeanBins <- bind_cols(ABins,AltBins)



AlterPaleo <- Alternative %>%
  merge(AltMeanBins) 

#names(AlterPaleo) <- c("Taxon", "CL", "extraCL", "PL", "size", "estimated", "Age", "Island", "Continent", "Genus", "bin", "tt")

paleoAlternativeGenus <- AlterPaleo %>%
  select(CL, tt=bins, Genus) %>%
  filter( !is.na(CL)) %>%
  group_by(Genus, tt) %>%
  summarise(meanCL = mean(CL), varCL=var(CL), n=n())


paleoAlternativeGenus[is.na(paleoAlternativeGenus)] <- 0

GenusAlter <- paleoAlternativeGenus %>%
  ungroup() %>%
  select(meanCL, tt) %>%
  group_by(tt) %>%
  summarise(mm = mean(meanCL), vv=var(meanCL), nn=n()) %>%
  arrange(tt)

GenusAlter[is.na(GenusAlter)] <- 0  


kable(GenusAlter, caption="paleoTS object, equal time bins, genera)")

GenusAlt <-as.paleoTS(GenusAlter$mm, GenusAlter$vv, GenusAlter$nn, GenusAlter$tt, MM = NULL, genpars = NULL, label = "Testudinidae body size evolution mode", reset.time=TRUE)

plot(GenusAlt)


```

```{r "Model-fitting, different time bins, genera" , echo=FALSE, message=FALSE, include=FALSE}

GenusAltFit <- (fit3models(GenusAlt, silent=FALSE, method="AD", pool=FALSE))
```



```{r "Model-fitting, different time bins, genera, table", echo=FALSE}

kable(GenusAltFit, caption="Model-fitting results for testudinidae, equal time bins, genera")

```

\newpage
## larger equal bins
### genera (larger equal bins)


```{r "Play around with larger time bins, generic level", echo=FALSE, fig.cap="Larger equal bins, genera"}

Alternative <- testRatio

Alternative$bin <- cut(Alternative$Age, c(0:50.000))


AltBins <- data.frame(c( (0+1)/2,  (1+2)/2,  (2+3)/2, (3+4)/2,  (4+5)/2, (5+6)/2,   
(6+7)/2,  (7+8)/2,  (8+9)/2,  (9+10)/2,   
(10+11)/2,  (11+12)/2, (12+13)/2,  (13+14)/2,  (14+15)/2,  (15+16)/2,  (16+17)/2,   (17+18)/2,
(18+19)/2,  (19+20)/2,  (23+24)/2,   (25+26)/2,  (32+33)/2,   (33+34)/2,    (49+50)/2      ))

names(AltBins) <- "bins"



ABins <- Alternative %>%
  select(bin) %>%
  unique() %>%
  arrange(bin)


AltMeanBins <- bind_cols(ABins,AltBins)



AlterPaleo <- Alternative %>%
  merge(AltMeanBins) 

#names(AlterPaleo) <- c("Taxon", "CL", "extraCL", "PL", "size", "estimated", "Age", "Island", "Continent", "Genus", "bin", "tt")

paleoAlternativeGenus <- AlterPaleo %>%
  select(CL, tt=bins, Genus) %>%
  filter( !is.na(CL)) %>%
  group_by(Genus, tt) %>%
  summarise(meanCL = mean(CL), varCL=var(CL), n=n())


paleoAlternativeGenus[is.na(paleoAlternativeGenus)] <- 0

GenusAlter <- paleoAlternativeGenus %>%
  ungroup() %>%
  select(meanCL, tt) %>%
  group_by(tt) %>%
  summarise(mm = mean(meanCL), vv=var(meanCL), nn=n()) %>%
  arrange(tt)

GenusAlter[is.na(GenusAlter)] <- 0  


kable(GenusAlter, caption="PaleoTS object, larger equal bins, genera")

GenusAlt <-as.paleoTS(GenusAlter$mm, GenusAlter$vv, GenusAlter$nn, GenusAlter$tt, MM = NULL, genpars = NULL, label = "Testudinidae body size evolution mode", reset.time=TRUE)

plot(GenusAlt)


```

```{r "Model-fitting, larger time bins, genus" , echo=FALSE, message=FALSE, include=FALSE}

GenusAltFit <- (fit3models(GenusAlt, silent=FALSE, method="AD", pool=FALSE))
```



```{r "Model-fitting, larger time bins, genus, table", echo=FALSE}

kable(GenusAltFit, caption="Model-fitting results for testudinidae, larger equal time bins, genera")

```


\newpage
## per continent
 
### Europe, smaller original bins (see Table 2), genera

```{r "paleoTS with different time bins, no bins, genera, Europe", echo=FALSE, fig.cap="Smaller original bins, genera, Europe"}

GenusMeanAll <- PleiPlioCL%>%
#  merge(BINS) %>%
#  filter(EpochBins != "Modern") %>%  # & EpochBins != "Upper Miocene"
  merge(Continents) %>%
  filter(Con == "Europe") %>%
  filter(!is.na(CL)) %>%
  group_by(MeanBins, Genus) %>%
  summarise(meanCL = mean(CL), n=n()) %>%
#  ungroup() %>%
#  group_by(EpochBins) %>%
  summarise(mm = mean(meanCL), nn=n(), vv=var(meanCL)) %>%
 # merge(BINS) %>%
  select(mm, nn, vv, tt=MeanBins)
  
GenusMeanAll[is.na(GenusMeanAll)]<-0



GenusPaleoAll <- GenusMeanAll

GenusPaleoAll[is.na(GenusPaleoAll)] <- 0

#kable(GenusPaleoAll)


paleoGenAll <-as.paleoTS(GenusPaleoAll$mm, GenusPaleoAll$vv, GenusPaleoAll$nn, GenusPaleoAll$tt, MM = NULL, genpars = NULL, label = "Testudinidae body size evolution mode", reset.time=TRUE)

plot(paleoGenAll)


```

```{r "Model-fitting, genera, all, no bins, Europe" , echo=FALSE, message=FALSE, include=FALSE}

PaleoGenAllFit <- (fit3models(paleoGenAll, silent=FALSE, method="AD", pool=FALSE))

#PaleoGenAllFit2 <- (fit4models(paleoGenAll, silent=FALSE, method="AD", pool=FALSE))

```



```{r "Model-fitting, genera, all, no bins, Europe, table", echo=FALSE}

kable(PaleoGenAllFit, caption="Model-fitting results for testudinidae, no bins, genera")

```



\newpage 
### Eurasia, smaller original bins (See Table 2), genera

```{r "paleoTS with different time bins, no bins, genera, Eurasia", echo=FALSE, fig.cap="Smaller original bins, genera, Eurasia"}

GenusMeanAll <- PleiPlioCL %>%
#  merge(BINS) %>%
#  filter(EpochBins != "Modern") %>%  # & EpochBins != "Upper Miocene"
  merge(Continents) %>%
  filter(Con == "Asia" | Con == "Europe") %>%
  filter(!is.na(CL)) %>%
  group_by(MeanBins, Genus) %>%
  summarise(meanCL = mean(CL), n=n()) %>%
#  ungroup() %>%
#  group_by(EpochBins) %>%
  summarise(mm = mean(meanCL), nn=n(), vv=var(meanCL)) %>%
 # merge(BINS) %>%
  select(mm, nn, vv, tt=MeanBins)
  
GenusMeanAll[is.na(GenusMeanAll)]<-0



sumTort <- read.csv("tortoises_summary.csv", sep=";", header=TRUE)
colnames(sumTort)[7] <- "meanCL"
colnames(sumTort)[8] <- "sdCL"


sumTortoisesAll <- sumTort %>%
#  dplyr::filter(Island == "y") %>%
  merge(Continents) %>%
  filter(Con== "Asia" | Con == "Europe") %>%
  mutate(Genus= as.character(Genus)) %>%
  mutate(MeanBins=(Mamin+Mamax)/2) %>%
  select(Genus, MeanBins,  meanCL, sdCL, n) #%>% #n,
 # bind_rows(GenusMeanAll)


SumTortAll <- sumTortoisesAll %>%
  group_by(Genus) %>%
  mutate(tt=MeanBins, vv=sdCL^2, nn=n, mm=meanCL) %>%
  dplyr::select(mm, nn, vv, tt, Genus) %>%
  mutate(nx = nn*mm) %>%
  mutate(mmall=sum(nx)/sum(nn)) %>%
  mutate(SD=sqrt(nx), d=mm-mmall) %>%
  mutate(nsd=((nx^2+d^2)*nn)) %>%
  mutate(varall=sum(nsd)/sum(nn), n=sum(nn)) %>%
  dplyr::select(mm=mmall, vv=varall, nn=n, tt, Genus) %>%
  unique() %>%
  dplyr::select(CL=mm, n=nn, var=vv, tt, Genus)


modernMeanGenusAll <- SumTortAll %>%
  ungroup() %>%
  select(CL, n, tt) %>%
  filter( !is.na(CL)) %>%
  group_by(tt) %>%
  summarise(mm = mean(CL), vv=var(CL), nn=n())

#bis hier alle modernen taxa zusammengefasst (summarised and MFN_testudinidae)


GenusPaleoAll <- modernMeanGenusAll %>%
  bind_rows(GenusMeanAll)%>%
  arrange(tt)

GenusPaleoAll[is.na(GenusPaleoAll)] <- 0

#kable(GenusPaleoAll)


paleoGenAll <-as.paleoTS(GenusPaleoAll$mm, GenusPaleoAll$vv, GenusPaleoAll$nn, GenusPaleoAll$tt, MM = NULL, genpars = NULL, label = "Testudinidae body size evolution mode", reset.time=TRUE)

plot(paleoGenAll)


```

```{r "Model-fitting, genera, all, no bins, Eurasia" , echo=FALSE, message=FALSE, include=FALSE}

PaleoGenAllFit <- (fit3models(paleoGenAll, silent=FALSE, method="AD", pool=FALSE))

#PaleoGenAllFit2 <- (fit9models(paleoGenAll, silent=FALSE, method="AD", pool=FALSE))

```



```{r "Model-fitting, genera, all, no bins, Eurasia, table", echo=FALSE}

kable(PaleoGenAllFit, caption="Model-fitting results for testudinidae, no bins, genera")

```

