---
title: "MAthesis"
#author: "Julia Joos"
#date: "14 August 2017"
output: 
  pdf_document:
   # toc: true
   # toc_depth: 2
    number_sections: true
    fig_caption: true
    df_print: kable
   # fontsize: 12pt
   # geometry: margin=1in
   # documentclass: scrartcl
   # mainfont: Helvetica
    citation_package: natbib
    keep_tex: true

header-includes:
    - \usepackage{setspace}
    - \doublespacing
    
# includes:
#       before_body: titlepage.tex
    
---

```{r "setup", include=FALSE}
require("knitr")
#opts_knit$set(root.dir = "//naturkundemuseum-berlin.de/MuseumDFSRoot/Benutzer/Julia.Joos/Eigene Dateien/MA")
## on laptop: 
opts_knit$set(root.dir = "C:/Users/Jule/Documents/Uni/MA")
library(dplyr)
library(ggplot2)
library(paleoTS)
```




```{r "import data set", echo=FALSE}
#### Data basis - import data set from Excel ####
setwd("C:/Users/Jule/Documents/Uni/MA")
tidyCL<-read.csv("tortoises_tidy.csv", sep=";", header=TRUE)

#####prepare for analysis (fix column names in .csv-file after converting table from excel####
colnames(tidyCL)[6] <- "MAmin"
colnames(tidyCL)[7] <- "Mamax"
colnames(tidyCL)[17] <- "CL"
colnames(tidyCL)[18] <- "PL"
colnames(tidyCL)[21] <- "estimated"

tidyCL <-  tidyCL %>%
  mutate(Age= ((MAmin)+(as.numeric(Mamax)))/2)

####### import extant data ####
extant <- read.csv("MFN_testudinidae.csv", sep=";", header=TRUE, dec=".", na.strings = "NA", stringsAsFactors=FALSE)  # file: MFN_testudinidae.csv

colnames(extant)[5] <- "SCL"
colnames(extant)[11] <- "PL"
colnames(extant)[12] <- "PLmid"



Extant <- extant %>%
  mutate(CL = SCL * 10, PL=PL*10, PLmid=PLmid*10) 
```


```{r "estimating CL from PL", echo=FALSE, results='asis'}
CLPLtidy <- tidyCL %>%
  filter(!is.na(CL) & !is.na(PL)) %>%
  dplyr::select(Taxon, CL, PL, size, Age, Island, Continent, Genus) %>%
  mutate(ratio=CL/PL) %>%
  group_by(Taxon) %>% #to show ratios per Taxon, leave out to get a total ratio
  summarise(meanRatio=round(mean(ratio),2), sdRatio=round(sd(ratio),2), n=n(), min=min(ratio), max=max(ratio))

CLPLextant <- Extant %>%
  filter(!is.na(CL) & !is.na(PL)) %>%
  dplyr::select(Taxon=Species, CL, PL, PLmid, Island, Continent, Genus) %>%
  mutate(ratio=CL/PL, ratioMid=CL/PLmid)
  #group_by(Taxon) %>% #to show ratios per Taxon, leave out to get a total ratio
  #  summarise(meanRatio=round(mean(ratio),2), sdRatio=round(sd(ratio),2), n=n(), min=min(ratio), max=max(ratio))

Ratio <- CLPLextant %>%
  summarise(meanRatio=round(mean(ratio),2), sdRatio=round(sd(ratio),2), n=n(), min=min(ratio), max=max(ratio))

RatioSpecies <- CLPLextant %>%
  group_by(Taxon) %>% #to show ratios per Taxon, leave out to get a total ratio
  summarise(meanRatio=round(mean(ratio),2), sdRatio=round(sd(ratio),2), n=n(), min=min(ratio), max=max(ratio))


fossil <- tidyCL %>%
  mutate(Taxon=as.character(Taxon), Island=as.character(Island), Continent=as.character(Continent), estimated=as.character(estimated), Genus=as.character(Genus)) %>%
  dplyr::select(Taxon, CL, PL, size, estimated, Age, Island, Continent, Genus)
  
modern <- Extant %>%
  dplyr::select(Taxon=Species, CL, PL, estimated, Age, Island, Continent, Genus)

All <- bind_rows(modern, fossil)

testRatio <- All %>% #tidyCL
  dplyr::select(Taxon, CL, PL, size, estimated, Age, Island, Continent, Genus) %>%
  mutate(extraCL = PL*Ratio$meanRatio) %>%
  dplyr::select(Taxon, CL, extraCL, PL, size, estimated, Age, Island, Continent, Genus)

testRatio$CL[is.na(testRatio$CL)] <- testRatio$extraCL[is.na(testRatio$CL)]
```


```{r "Bin data", echo=FALSE}

PleiPlio <- testRatio %>% #testRatio or tidyCL 
  filter(Age < 11.000)

PleiPlio$bin <- cut(PleiPlio$Age, c(0, 0.000001, 0.0117, 0.126, 0.781, 2.588, 3.6, 5.332, 11.608))

EpochBins <- as.vector(c("Modern", "Holocene", "Upper Pleistocene", "Middle Pleistocene", "Lower Pleistocene", "Upper Pliocene", "Lower Pliocene", "Upper Miocene"))




MeanBins <- as.vector(c((0+0.000001)/2, (0+0.0117)/2, (0.0117+0.126)/2, (0.126+0.781)/2, (0.781+2.588)/2,(2.588+3.6)/2, (3.6+5.332)/2, (5.332+11.608)/2))

Bins <- PleiPlio %>%
  select(bin) %>%
  group_by(bin) %>%
  summarise(nIndividuals=n())

BinsSpecies <- PleiPlio %>%
  select(bin, Taxon) %>%
  group_by(bin, Taxon) %>%
  summarise(nSpecies=n()) %>%
  summarise(nSpecies=n())


BinsGenera <- PleiPlio %>%
  select(bin, Genus) %>%
  group_by(bin, Genus) %>%
  summarise(nGenera=n()) %>%
  summarise(nGenera=n())

bin <- as.vector(unique(Bins$bin))



BINS <- data.frame(bin, EpochBins, MeanBins) %>%
  merge(Bins) %>%
  merge(BinsSpecies) %>%
  merge(BinsGenera) %>%
  arrange(MeanBins)

BINS$EpochBins = factor(BINS$EpochBins, levels=c("Modern", "Holocene", "Upper Pleistocene",
                                                             "Middle Pleistocene", "Lower Pleistocene",
                                                             "Upper Pliocene", "Lower Pliocene",
                                                             "Upper Miocene"))

#BINS <- read.table("timebins.txt", sep="\t", header=TRUE)

#kable(Bins, caption="Time bins with corresponding sample sizes (individuals)")

kable(BINS, caption="Time bins with age range, epoch name, mean age and corresponding sample sizes (on individual, species and genus level)")

PleiPlioCL <- PleiPlio %>%
  merge(BINS)

SUM <- PleiPlioCL %>%
  group_by(MeanBins) %>%
  summarise(n=n())
```


```{r "Map fossil occurrences", echo=FALSE, fig.cap="Map displaying all fossil occurrences of testudinids, with color indicating whether relevant literature was available (black if not) and if it was, whether body size data was available or not (yes and no, respectively)."}

CL <- tidyCL %>%
#  filter(Age < 11.000) %>%
  dplyr::select(Locality, Genus, Taxon, Latitude, Longitude, Country, Age) %>%
  group_by(Locality) %>%
  mutate(count= n())


FossilOccurrences<-read.csv("tortoises_occurrences.csv", sep=";", header=TRUE)

FossOcc <- FossilOccurrences %>%
  mutate(Age= (as.numeric(as.character(MA.min)) + as.numeric(as.character(Ma.max)))/2) %>%
  select(Locality, Country, Latitude, Longitude, Age, Genus, Taxon, Clavailability) %>%
#  merge(tidyCL) %>%
  group_by(Locality) %>%
  mutate(count= n(), Longitude=as.numeric(as.character(Longitude)))


#na <- FossOcc[!complete.cases(FossOcc),]

# OccMap <-  merge(Map, FossOcc, by="Locality", incomparables=TRUE) %>%
#  distinct()
  
# unique(CL$Locality) #wie viele localities
#length(unique(FossOcc$Locality[which(FossOcc$Clavailability == "yes")]))


mapWorld <- borders("world", colour="azure3", fill="azure3") # create a layer of borders, run line at the beginning (before loading plotly)

cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")


mapOc <- FossOcc %>%
  ggplot(aes(Longitude, Latitude)) +# borders("world", ylim = c(-60, 90)) +
  mapWorld +
  theme_classic() +
  geom_point(data=FossOcc, (aes(as.numeric(Longitude), Latitude, colour=FossOcc$Clavailability))) +
  geom_point(data=CL,(aes(Longitude, Latitude, colour="yes"))) +
  scale_colour_manual(values=cbbPalette)

#unique(FossOcc$CL)

mapOc

# require(plotly)
# ggplotly(mapOc)


######

# length(CL$Locality)
# length(unique(CL$Locality)) #wie viele localities
# length(unique(FossOcc$Locality[which(FossOcc$Clavailability == "yes")]))



```


```{r "Map body size data set", echo=FALSE, fig.cap="Map displaying all localities for which body size data for testudinids was available in the literature. Size of points denotes sample size, color denotes approximate age."}


tidyCL$bin <- cut(tidyCL$Age, c(0, 0.000001, 0.0117, 0.126, 0.781, 2.588, 3.6, 5.332, 11.608,33.9, 56.000))


Bins <- tidyCL %>%
  select(bin) %>%
  group_by(bin) %>%
  summarise(n=n())


bin <- as.vector(unique(Bins$bin))
EpochBins <- as.vector(c("Holocene", "Upper Pleistocene", "Middle Pleistocene", "Lower Pleistocene", "Upper Pliocene", "Lower Pliocene", "Upper Miocene", "Oligocene", "Eocene"))

MeanBins <- as.vector(c((0+0.0117)/2, (0.0117+0.126)/2, (0.126+0.781)/2, (0.781+2.588)/2,(2.588+3.6)/2, (3.6+5.332)/2, (5.332+11.608)/2, (11.608+33.9)/2, (33.9+56.0)/2 ))


bin <- as.vector(unique(Bins$bin))

BINS <- data.frame(bin, EpochBins, MeanBins) #


#BINS <- data.frame(bin, EpochBins, MeanBins) #
#BINS <- read.table("timebins.txt", sep="\t", header=TRUE)


MapCL <- tidyCL %>%
  merge(BINS) %>%
#  filter(Age < 11.000) %>%
#  filter(Latitude != "-") %>%
  dplyr::select(Genus, Taxon, Latitude, Longitude, Country, CL, PL, MeanBins, size) %>%
  group_by(Latitude) %>%
  mutate(count= n()) 


mapWorld <- borders("world", colour="azure3", fill="azure3") # create a layer of borders, run line at the beginning (before loading plotly)


mapCL <- MapCL %>%
  ggplot(aes(Longitude, Latitude)) +# borders("world", ylim = c(-60, 90)) +
  mapWorld +
  theme_classic() +
  geom_point(aes(Longitude, Latitude,size=count, colour=MeanBins)) +  #colour=size, 
  scale_colour_gradientn(colors=c("orange", "red", "purple", "blue", "green"))



mapCL


```




```{r "Get overview over sample sizes", echo=FALSE, include=FALSE}
OverviewSpecies <- PleiPlioCL %>%
  filter(EpochBins != "Modern") %>%
  group_by(EpochBins, Taxon) %>%
  summarise(n=n())

OverviewSpeciesFossil <- PleiPlioCL %>%
  filter(EpochBins != "Modern") %>%
  group_by(Taxon) %>%
  summarise(n=n())

#write.table(OverviewSpecies,file="OverviewSpeciesSampleSize.txt", sep="\t", row.names = FALSE) 


OverviewGeneraTime <- PleiPlioCL %>%
  group_by(EpochBins, Genus) %>%
  summarise(n=n())

OverviewGenera <- PleiPlioCL %>%
  group_by(Genus) %>%
  summarise(n=n())

```







```{r "Histograms of body size data", echo=FALSE, fig.cap="Distribution of body site data per time bin"}


# HistCLModern <- PleiPlioCL %>%
#   filter( !is.na(CL)) %>%
#   filter(EpochBins == "Modern") %>%
#   ggplot(aes(CL)) + geom_histogram(binwidth=10, col="black", fill="gray") + facet_wrap(~EpochBins, scales="free_x")+
#   theme_classic()+ 
#   #theme_gray() +
#   theme(panel.border = element_rect(colour = "black", fill=NA))

HistCL <- PleiPlioCL %>%
  filter( !is.na(CL)) %>%
#  filter(EpochBins != "Modern") %>%
  ggplot(aes(CL)) + geom_histogram(binwidth=10, col="black", fill="gray") + facet_wrap(~EpochBins, scales="free_x")+
  theme_classic()+ 
  #theme_gray() +
  theme(panel.border = element_rect(colour = "black", fill=NA))

#HistCLModern
HistCL  

```


```{r "Species Accumulation Curve", echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Species Accumulation Curve of fossil species per Locality and reference"}

vegan <- tidyCL %>%
  dplyr::select(Locality, Taxon) %>%
  group_by(Locality, Taxon) %>%
  summarise(n=n()) %>%
  tidyr::spread(Taxon, n, fill=0)

library(vegan)

#par(mfrow = c(2, 1))

vegan=vegan[,-1]
vegansp=specaccum(vegan,method="rarefaction", permutations=1000)
plot(vegansp,xlab="Ind",ylab="Richness", xvar="individuals", ci.type="line", ci.lty=2, ci.col="grey", col="deepskyblue4", lwd=2, main="Fossil species, CL, per Locality" )



vegan <- tidyCL %>%
  dplyr::select(Reference, Taxon) %>%
  group_by(Reference, Taxon) %>%
  summarise(n=n()) %>%
  tidyr::spread(Taxon, n, fill=0)

library(vegan)

vegan=vegan[,-1]
vegansp=specaccum(vegan,method="rarefaction", permutations=1000)
plot(vegansp,xlab="Ind",ylab="Richness", xvar="individuals", ci.type="line", ci.lty=2, ci.col="grey", col="deepskyblue4", lwd=2, main="Fossil species, CL, per Reference" )


```



```{r "Species Accumulation Curve with Genera", echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Species Accumulation Curve of fossil genera per Locality and reference"}

vegan <- tidyCL %>%
  dplyr::select(Locality, Genus) %>%
  group_by(Locality, Genus) %>%
  summarise(n=n()) %>%
  tidyr::spread(Genus, n, fill=0)


#par(mfrow = c(2, 1))


vegan=vegan[,-1]
vegansp=specaccum(vegan,method="rarefaction", permutations=1000)
plot(vegansp,xlab="Ind",ylab="Richness", xvar="individuals", ci.type="line", ci.lty=2, ci.col="grey", col="deepskyblue4", lwd=2, main="Fossil genera, CL, per Locality" )


vegan <- tidyCL %>%
  dplyr::select(Reference, Genus) %>%
  group_by(Reference, Genus) %>%
  summarise(n=n()) %>%
  tidyr::spread(Genus, n, fill=0)

library(vegan)

vegan=vegan[,-1]
vegansp=specaccum(vegan,method="rarefaction", permutations=1000)
plot(vegansp,xlab="Ind",ylab="Richness", xvar="individuals", ci.type="line", ci.lty=2, ci.col="grey", col="deepskyblue4", lwd=2, main="Fossil genera, CL, per Reference")


```


```{r "Boxplots of each genus per time bin", echo=FALSE, warning=FALSE, fig.cap="Boxplots of each genus per time bin, for colors see Fig. 4."}
PleiPlioCL$EpochBins = factor(PleiPlioCL$EpochBins, levels=c("Modern", "Holocene", "Upper Pleistocene",
                                                             "Middle Pleistocene", "Lower Pleistocene",
                                                             "Upper Pliocene", "Lower Pliocene",
                                                             "Upper Miocene"))

IndGenera <- PleiPlioCL %>%
  group_by(EpochBins, Genus) %>%
  filter(!is.na(CL)) %>%
  summarise(GenusMean=mean(CL), GenusSD=sd(CL), n=n()) 


BinGenera <- PleiPlioCL %>%
  group_by(EpochBins, Genus) %>%
  filter(!is.na(CL)) %>%
  summarise(GenusMean=mean(CL), GenusSD=sd(CL), n=n()) %>%
#  summarise(TimeBinMean=mean(GenusMean), TimeBinSD=sd(GenusMean), n=n()) %>%
  ungroup() %>%#, CL=CL
  ggplot(aes(EpochBins, GenusMean)) + #, colour=Genus
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  stat_boxplot(geom ='errorbar', width = 0.2) +
  geom_boxplot() +
  geom_pointrange(data=IndGenera, position="jitter", aes(x=EpochBins, y=GenusMean, colour=Genus, ymin = GenusMean-GenusSD, ymax = GenusMean+GenusSD)) +
  guides(color="none")
#  geom_text(data=IndGenera,aes(label=Genus, x=EpochBins, y=GenusMean))
#  geom_errorbar(data=IndGenera, position="jitter",aes(ymin = GenusMean-GenusSD, ymax = GenusMean+GenusSD, width=0.2, colour=Genus))#+
  #scale_color_brewer(palette="Set1")
  #scale_colour_gradientn(colours=rainbow(4))

BinGenera

```


```{r "Separate genera per time bin", echo=FALSE, warning=FALSE,  fig.cap="Mean body size and standard deviation per genus in each time bin"}
BinGeneraFacet <- PleiPlioCL %>%
  group_by(EpochBins, Genus) %>%
  filter(!is.na(CL)) %>%
  summarise(GenusMean=mean(CL), GenusSD=sd(CL), n=n()) %>%
#  summarise(TimeBinMean=mean(GenusMean), TimeBinSD=sd(GenusMean), n=n()) %>%
  ungroup() %>%#, CL=CL
  ggplot(aes(EpochBins, GenusMean)) + #, colour=Genus
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
#  stat_boxplot(geom ='errorbar', width = 0.2) +
#  geom_boxplot() +
  geom_point(data=IndGenera,  aes(x=Genus, y=GenusMean, colour=Genus)) +
  geom_pointrange(data=IndGenera, aes(x=Genus, y=GenusMean, colour=Genus, ymin = GenusMean-GenusSD, ymax = GenusMean+GenusSD)) +
#  geom_boxplot(data=IndGenera,  aes(x=Genus, y=GenusMean, colour=Genus))+
#  geom_text(data=IndGenera,aes(label=Genus, x=EpochBins, y=GenusMean)) +
  facet_wrap(~EpochBins, scales="free_x") +
  guides(colour="none")#+
#  stat_boxplot(data=BinGenera, geom ='errorbar', width = 0.2) +
#  geom_boxplot(data=BinGenera) 
#  geom_errorbar(data=IndGenera, position="jitter",aes(ymin = GenusMean-GenusSD, ymax = GenusMean+GenusSD, width=0.2, colour=Genus))#+
  #scale_color_brewer(palette="Set1")
  #scale_colour_gradientn(colours=rainbow(4))

BinGeneraFacet
```


\newpage

# paleoTS analysis

## all (continental and insular)

### individuals (all)

```{r "paleoTS, individuals, inluding island species", echo=FALSE}
PPCL <- PleiPlioCL %>%
  dplyr::select(CL, MeanBins) %>%
  dplyr::filter(CL != "NA") %>%
  mutate(tt= MeanBins) %>% # create mean age
  group_by(tt) %>% #create time bins
  summarise(mm=mean(CL), vv=var(CL), nn=n()) #create means etc. for each time bin 

PPCL[is.na(PPCL)]<-0 #subset NAs with O for n=1


sumTort <- read.csv("tortoises_summary.csv", sep=";", header=TRUE)

SumTort <- sumTort %>%
  mutate(tt=(Mamin+Mamax)/2, vv=sdCLmm^2, nn=n, mm=meanCLmm) %>%
  dplyr::select(mm, nn, vv, tt)
 
PPESCL <- bind_rows(SumTort,PPCL) #ExTort, 

combiCL <- PPESCL %>%
  dplyr::filter(tt==0.0000005) %>%   #maybe find a way to automatically filter double tt values...
  mutate(nx = nn*mm) %>%
  mutate(mmall=sum(nx)/sum(nn)) %>%
  mutate(SD=sqrt(nx), d=mm-mmall) %>%
  mutate(nsd=((nx^2+d^2)*nn)) %>%
  mutate(varall=sum(nsd)/sum(nn), n=sum(nn)) %>%
  dplyr::select(mm=mmall, vv=varall, nn=n, tt) %>%
  unique() %>%
  dplyr::select(mm, nn, vv, tt)


PPCL <- PPESCL %>%
  filter(tt !=0.0000005) %>%
  bind_rows(combiCL)%>%
  arrange(tt)

```



```{r "paleoTS object", echo=FALSE, include=FALSE}

kable(PPCL, caption="paleoTS object (mm= mean CL, nn = sample size, vv = variance (CL), tt = Age)")
```



```{r "paleoTS plot", echo=FALSE, fig.cap="individuals, including island species"}
library(paleoTS)
paleoPPCL <-as.paleoTS(PPCL$mm, PPCL$vv, PPCL$nn, PPCL$tt, MM = NULL, genpars = NULL, label = "Testudinidae body size evolution mode, Holocene-Upper Miocene, CL, including Island species")
#paleoPPCL
plot(paleoPPCL)

```



```{r "Model-fitting, individuals" , echo=FALSE, message=FALSE, include=FALSE}

PaleoPCCLFit <- (fit3models(paleoPPCL, silent=FALSE, method="AD", pool=FALSE))
```



```{r "Model-fitting, individuals, table", echo=FALSE}

kable(PaleoPCCLFit, caption="Model-fitting results for testudinidae, individuals, including island species")

```


\newpage

### species (all)

```{r "paleoTS plot with species mean, including island species", echo=FALSE, fig.cap="paleoTS plot with species mean, including island species"}
SpeciesMean <- PleiPlioCL %>%
  filter(EpochBins != "Modern") %>%
  group_by(EpochBins, Taxon, Age) %>%
  summarise(meanCL = mean(CL), n=n()) %>%
  merge(BINS)

SpeciesModern <- PleiPlioCL %>%
  filter(EpochBins == "Modern") %>%
  group_by(Taxon) %>%
  summarise(meanCL=mean(CL), sdCL=sd(CL), n=n(), Age=mean(Age), EpochBins=unique(EpochBins), MeanBins=unique(MeanBins))


SpeciesModern[is.na(SpeciesModern)]<-0 #subset NAs with O for n=1



SpeciesOverview <- SpeciesMean %>%
#  filter(EpochBins != "Modern") %>%
  filter( !is.na(meanCL)) %>%
  group_by(EpochBins) %>%
  summarise(meanSpeciesCL = mean(meanCL), nSpecies=n(), MeanBins = mean(MeanBins)) %>%
  arrange(MeanBins) 

#kable(SpeciesOverview)




SpeciesPaleo <- SpeciesMean %>%
  dplyr::select(meanCL, MeanBins) %>%
  dplyr::filter(meanCL != "NA") %>%
  mutate(tt= MeanBins) %>% # create mean age
  group_by(tt) %>% #create time bins
  summarise(mm=mean(meanCL), vv=var(meanCL), nn=n()) #create means etc. for each time bin 

#SpeciesPaleo[is.na(SpeciesPaleo)]<-0 #subset NAs with O for n=1


sumTort <- read.csv("tortoises_summary.csv", sep=";", header=TRUE)
#colnames(sumTort)[1] <- "Taxon"
colnames(sumTort)[6] <- "meanCL"
colnames(sumTort)[7] <- "sdCL"
#sumTort$EpochBins <- "Modern"
#sumTort$MeanBins <- 0.0000005


sumTortoises <- sumTort %>%
  mutate(Taxon= as.character(Species)) %>%
  mutate(MeanBins=(Mamin+Mamax)/2) %>%
  select(Taxon, MeanBins,  meanCL, sdCL) %>% #n,
  bind_rows(SpeciesModern)


SumTort <- sumTortoises %>%
  group_by(Taxon) %>%
  mutate(tt=MeanBins, vv=sdCL^2, nn=n, mm=meanCL) %>%
  dplyr::select(mm, nn, vv, tt, Taxon) %>%
  mutate(nx = nn*mm) %>%
  mutate(mmall=sum(nx)/sum(nn)) %>%
  mutate(SD=sqrt(nx), d=mm-mmall) %>%
  mutate(nsd=((nx^2+d^2)*nn)) %>%
  mutate(varall=sum(nsd)/sum(nn), n=sum(nn)) %>%
  dplyr::select(mm=mmall, vv=varall, nn=n, tt, Taxon) %>%
  unique() %>%
  dplyr::select(CL=mm, n=nn, var=vv, tt, Taxon)

# write.table(SumTort,file="SumTortModern.txt", sep="\t", row.names = FALSE) 

#boxplot(SumTort$CL, caption="Body size distribution in time bin 'modern'.")


modernMean <- SumTort %>%
  ungroup() %>%
  select(CL, n, tt) %>%
  filter( !is.na(CL)) %>%
  group_by(tt) %>%
  summarise(mm = mean(CL), vv=var(CL), nn=n())

#bis hier alle modernen taxa zusammengefasst (summarised and MFN_testudinidae)



SpecPaleo <- bind_rows(modernMean,SpeciesPaleo) %>%
  arrange(tt)


#kable(SpecPaleo)


paleoSpec <-as.paleoTS(SpecPaleo$mm, SpecPaleo$vv, SpecPaleo$nn, SpecPaleo$tt, MM = NULL, genpars = NULL, label = "Testudinidae body size evolution mode, Holocene-Upper Miocene, species mean CL", reset.time=TRUE)

plot(paleoSpec)
```




```{r "Model-fitting, species" , echo=FALSE, message=FALSE, include=FALSE}

PaleoSpecFit <- (fit3models(paleoSpec, silent=FALSE, method="AD", pool=FALSE))
```



```{r "Model-fitting, species, table", echo=FALSE}

kable(PaleoSpecFit, caption="Model-fitting results for testudinidae, species, including island species")

```


\newpage

### genera (all)

```{r "paleoTS plot with genus mean, including island species", echo=FALSE, fig.cap="paleoTS plot with genus mean, including island species"}
GenusMean <- PleiPlioCL %>%
  filter(EpochBins != "Modern") %>%
  filter(!is.na(CL)) %>%
  group_by(EpochBins, Genus) %>%
  summarise(meanCL = mean(CL), n=n()) %>%
#  ungroup() %>%
#  group_by(EpochBins) %>%
  summarise(mm = mean(meanCL), nn=n(), vv=var(meanCL)) %>%
  merge(BINS) %>%
  select(mm, nn, vv, tt=MeanBins)
  
#na <- PleiPlioCL[!complete.cases(PleiPlioCL),]



GenusModern <- PleiPlioCL %>%
  filter(EpochBins == "Modern") %>%
  filter(!is.na(CL)) %>%
  group_by(Genus) %>%
  summarise(meanCL=mean(CL), sdCL=sd(CL), n=n(), Age=mean(Age), EpochBins=unique(EpochBins), MeanBins=unique(MeanBins)) #%>%
#  summarise(mm=mean(meanCL), nn=n(), vv=var(meanCL), tt=mean(MeanBins))



sumTort <- read.csv("tortoises_summary.csv", sep=";", header=TRUE)
#colnames(sumTort)[1] <- "Taxon"
colnames(sumTort)[7] <- "meanCL"
colnames(sumTort)[8] <- "sdCL"
#sumTort$EpochBins <- "Modern"
#sumTort$MeanBins <- 0.0000005


sumTortoises <- sumTort %>%
  mutate(Genus= as.character(Genus)) %>%
  mutate(MeanBins=(Mamin+Mamax)/2) %>%
  select(Genus, MeanBins,  meanCL, sdCL, n) %>% #n,
  bind_rows(GenusModern)


SumTort <- sumTortoises %>%
  group_by(Genus) %>%
  mutate(tt=MeanBins, vv=sdCL^2, nn=n, mm=meanCL) %>%
  dplyr::select(mm, nn, vv, tt, Genus) %>%
  mutate(nx = nn*mm) %>%
  mutate(mmall=sum(nx)/sum(nn)) %>%
  mutate(SD=sqrt(nx), d=mm-mmall) %>%
  mutate(nsd=((nx^2+d^2)*nn)) %>%
  mutate(varall=sum(nsd)/sum(nn), n=sum(nn)) %>%
  dplyr::select(mm=mmall, vv=varall, nn=n, tt, Genus) %>%
  unique() %>%
  dplyr::select(CL=mm, n=nn, var=vv, tt, Genus)

#write.table(SumTort,file="SumTortModernGenus.txt", sep="\t", row.names = FALSE)

#kable(SumTort, caption="Overview over body size means per time bin on genus level.")

#boxplot(SumTort$CL, caption="Body size distribution in time bin 'modern'.")


modernMeanGenus <- SumTort %>%
  ungroup() %>%
  select(CL, n, tt) %>%
  filter( !is.na(CL)) %>%
  group_by(tt) %>%
  summarise(mm = mean(CL), vv=var(CL), nn=n())

#bis hier alle modernen taxa zusammengefasst (summarised and MFN_testudinidae)


GenusPaleo <- modernMeanGenus %>%
  bind_rows(GenusMean)%>%
  arrange(tt)


#kable(GenusPaleo)


paleoGen <-as.paleoTS(GenusPaleo$mm, GenusPaleo$vv, GenusPaleo$nn, GenusPaleo$tt, MM = NULL, genpars = NULL, label = "Testudinidae body size evolution mode, Holocene-Upper Miocene, genus mean CL", reset.time=TRUE)

plot(paleoGen)

```



```{r "Model-fitting, genera" , echo=FALSE, message=FALSE, include=FALSE}

PaleoGenFit <- (fit3models(paleoGen, silent=FALSE, method="AD", pool=FALSE))
```



```{r "Model-fitting, genera, table", echo=FALSE}

kable(PaleoGenFit, caption="Model-fitting results for testudinidae, genera, including island species")

```




\newpage

## continental (excluding insular species)

### individuals (continental)

```{r "paleoTS, individuals, exluding island species", echo=FALSE, message=FALSE, fig.cap="individuals, excluding island species"}
PPCLI <- PleiPlioCL %>%
  dplyr::select(CL, MeanBins, Island) %>%
  dplyr::filter(CL != "NA" & Island == "n") %>%
 # dplyr::filter(PleiPlioCL$Island == "n") %>%
  mutate(tt= MeanBins) %>% # create mean age
  group_by(tt) %>% #create time bins
  summarise(mm=mean(CL), vv=var(CL), nn=n()) #create means etc. for each time bin 

#PPCLI[is.na(PPCLI)]<-0 #subset NAs with O for n=1


sumTort <- read.csv("tortoises_summary.csv", sep=";", header=TRUE)

SumTort <- sumTort %>%
  dplyr::filter(Island == "n") %>%
  mutate(tt=(Mamin+Mamax)/2, vv=sdCLmm^2, nn=n, mm=meanCLmm) %>%
  dplyr::select(mm, nn, vv, tt)
 
PPESCLI <- bind_rows(SumTort,PPCLI) #ExTort, 

combiCLI <- PPESCLI %>%
  dplyr::filter(tt==0.0000005) %>%   #maybe find a way to automatically filter double tt values...
  mutate(nx = nn*mm) %>%
  mutate(mmall=sum(nx)/sum(nn)) %>%
  mutate(SD=sqrt(nx), d=mm-mmall) %>%
  mutate(nsd=((nx^2+d^2)*nn)) %>%
  mutate(varall=sum(nsd)/sum(nn), n=sum(nn)) %>%
  dplyr::select(mm=mmall, vv=varall, nn=n, tt) %>%
  unique() %>%
  dplyr::select(mm, nn, vv, tt)


PPCLIs <- PPESCLI %>%
  filter(tt !=0.0000005) %>%
  bind_rows(combiCLI)%>%
  arrange(tt)



#kable(PPCLIs, caption="paleoTS object (mm= mean CL, nn = sample size, vv = variance (CL), tt = Age)")


library(paleoTS)
paleoPPCLIs <-as.paleoTS(PPCLIs$mm, PPCLIs$vv, PPCLIs$nn, PPCLIs$tt, MM = NULL, genpars = NULL, label = "Testudinidae body size evolution mode, Holocene-Upper Miocene, CL, including Island species")
#paleoPPCL
plot(paleoPPCLIs)
```




```{r "Model-fitting, individuals, continental" , echo=FALSE, message=FALSE, include=FALSE}

PaleoPCCLIsFit <- (fit3models(paleoPPCLIs, silent=FALSE, method="AD", pool=FALSE))
```



```{r "Model-fitting, individuals, continental, table", echo=FALSE}

kable(PaleoPCCLIsFit, caption="Model-fitting results for testudinidae, individuals, excluding island species")

```



\newpage


### species (continental)

```{r "paleoTS plot with species mean, excluding island species", echo=FALSE, message=FALSE, fig.cap="paleoTS plot with species mean, excluding island species"}
SpeciesMeanIs <- PleiPlioCL %>%
  filter(EpochBins != "Modern" & Island == "n") %>%
  group_by(EpochBins, Taxon, Age) %>%
  summarise(meanCL = mean(CL), n=n()) %>%
  merge(BINS)

SpeciesModernIs <- PleiPlioCL %>%
  filter(EpochBins == "Modern" & Island == "n") %>%
  group_by(Taxon) %>%
  summarise(meanCL=mean(CL), sdCL=sd(CL), n=n(), Age=mean(Age), EpochBins=unique(EpochBins), MeanBins=unique(MeanBins))


SpeciesModernIs[is.na(SpeciesModernIs)]<-0 #subset NAs with O for n=1



SpeciesOverviewIs <- SpeciesMeanIs %>%
#  filter(EpochBins != "Modern") %>%
  filter( !is.na(meanCL)) %>%
  group_by(EpochBins) %>%
  summarise(meanSpeciesCL = mean(meanCL), nSpecies=n(), MeanBins = mean(MeanBins)) %>%
  arrange(MeanBins) 

#kable(SpeciesOverviewIs)




SpeciesPaleoIs <- SpeciesMeanIs %>%
  dplyr::select(meanCL, MeanBins) %>%
  dplyr::filter(meanCL != "NA") %>%
  mutate(tt= MeanBins) %>% # create mean age
  group_by(tt) %>% #create time bins
  summarise(mm=mean(meanCL), vv=var(meanCL), nn=n()) #create means etc. for each time bin 

#SpeciesPaleo[is.na(SpeciesPaleo)]<-0 #subset NAs with O for n=1


sumTort <- read.csv("tortoises_summary.csv", sep=";", header=TRUE)
colnames(sumTort)[6] <- "meanCL"
colnames(sumTort)[7] <- "sdCL"


sumTortoises <- sumTort %>%
  dplyr::filter(Island == "n") %>%
  mutate(Taxon= as.character(Species)) %>%
  mutate(MeanBins=(Mamin+Mamax)/2) %>%
  select(Taxon, MeanBins,  meanCL, sdCL) %>% #n,
  bind_rows(SpeciesModernIs)


SumTort <- sumTortoises %>%
  group_by(Taxon) %>%
  mutate(tt=MeanBins, vv=sdCL^2, nn=n, mm=meanCL) %>%
  dplyr::select(mm, nn, vv, tt, Taxon) %>%
  mutate(nx = nn*mm) %>%
  mutate(mmall=sum(nx)/sum(nn)) %>%
  mutate(SD=sqrt(nx), d=mm-mmall) %>%
  mutate(nsd=((nx^2+d^2)*nn)) %>%
  mutate(varall=sum(nsd)/sum(nn), n=sum(nn)) %>%
  dplyr::select(mm=mmall, vv=varall, nn=n, tt, Taxon) %>%
  unique() %>%
  dplyr::select(CL=mm, n=nn, var=vv, tt, Taxon)

# write.table(SumTort,file="SumTortModern.txt", sep="\t", row.names = FALSE) 

#boxplot(SumTort$CL, caption="Body size distribution in time bin 'modern'.")


modernMeanIs <- SumTort %>%
  ungroup() %>%
  select(CL, n, tt) %>%
  filter( !is.na(CL)) %>%
  group_by(tt) %>%
  summarise(mm = mean(CL), vv=var(CL), nn=n())

#bis hier alle modernen taxa zusammengefasst (summarised and MFN_testudinidae)



SpecPaleoIs <- bind_rows(modernMeanIs,SpeciesPaleoIs) %>%
  arrange(tt)


#kable(SpecPaleoIs)


paleoSpecIs <-as.paleoTS(SpecPaleoIs$mm, SpecPaleoIs$vv, SpecPaleoIs$nn, SpecPaleoIs$tt, MM = NULL, genpars = NULL, label = "Testudinidae body size evolution mode, Holocene-Upper Miocene, species mean CL, continental", reset.time=TRUE)

plot(paleoSpecIs)
```



```{r "Model-fitting, species, continental" , echo=FALSE, message=FALSE, include=FALSE}

PaleoSpecIsFit <- (fit3models(paleoSpecIs, silent=FALSE, method="AD", pool=FALSE))
```



```{r "Model-fitting, species, continental, table", echo=FALSE}

kable(PaleoSpecIsFit, caption="Model-fitting results for testudinidae, species, excluding island species")

```




\newpage

### genera (continental)

```{r "paleoTS plot with genus mean, excluding island species", echo=FALSE, message=FALSE, fig.cap="paleoTS plot with genus mean, excluding island species"}
GenusMeanIs <- PleiPlioCL %>%
  filter(EpochBins != "Modern" & Island == "n") %>%
  filter(!is.na(CL)) %>%
  group_by(EpochBins, Genus) %>%
  summarise(meanCL = mean(CL), n=n()) %>%
#  ungroup() %>%
#  group_by(EpochBins) %>%
  summarise(mm = mean(meanCL), nn=n(), vv=var(meanCL)) %>%
  merge(BINS) %>%
  select(mm, nn, vv, tt=MeanBins)
  
#na <- PleiPlioCL[!complete.cases(PleiPlioCL),]



GenusModernIs <- PleiPlioCL %>%
  filter(EpochBins == "Modern" & Island == "n") %>%
  filter(!is.na(CL)) %>%
  group_by(Genus) %>%
  summarise(meanCL=mean(CL), sdCL=sd(CL), n=n(), Age=mean(Age), EpochBins=unique(EpochBins), MeanBins=unique(MeanBins)) #%>%
#  summarise(mm=mean(meanCL), nn=n(), vv=var(meanCL), tt=mean(MeanBins))



sumTort <- read.csv("tortoises_summary.csv", sep=";", header=TRUE)
colnames(sumTort)[7] <- "meanCL"
colnames(sumTort)[8] <- "sdCL"


sumTortoises <- sumTort %>%
  dplyr::filter(Island == "n") %>%
  mutate(Genus= as.character(Genus)) %>%
  mutate(MeanBins=(Mamin+Mamax)/2) %>%
  select(Genus, MeanBins,  meanCL, sdCL, n) %>% #n,
  bind_rows(GenusModernIs)


SumTort <- sumTortoises %>%
  group_by(Genus) %>%
  mutate(tt=MeanBins, vv=sdCL^2, nn=n, mm=meanCL) %>%
  dplyr::select(mm, nn, vv, tt, Genus) %>%
  mutate(nx = nn*mm) %>%
  mutate(mmall=sum(nx)/sum(nn)) %>%
  mutate(SD=sqrt(nx), d=mm-mmall) %>%
  mutate(nsd=((nx^2+d^2)*nn)) %>%
  mutate(varall=sum(nsd)/sum(nn), n=sum(nn)) %>%
  dplyr::select(mm=mmall, vv=varall, nn=n, tt, Genus) %>%
  unique() %>%
  dplyr::select(CL=mm, n=nn, var=vv, tt, Genus)

#write.table(SumTort,file="SumTortModernGenus.txt", sep="\t", row.names = FALSE)

#kable(SumTort, caption="Overview over body size means per time bin on genus level, excluding island species.")

#boxplot(SumTort$CL, caption="Body size distribution in time bin 'modern'.")


modernMeanGenusIs <- SumTort %>%
  ungroup() %>%
  select(CL, n, tt) %>%
  filter( !is.na(CL)) %>%
  group_by(tt) %>%
  summarise(mm = mean(CL), vv=var(CL), nn=n())

#bis hier alle modernen taxa zusammengefasst (summarised and MFN_testudinidae)


GenusPaleoIs <- modernMeanGenusIs %>%
  bind_rows(GenusMeanIs)%>%
  arrange(tt)


#kable(GenusPaleoIs)


paleoGenIs <-as.paleoTS(GenusPaleoIs$mm, GenusPaleoIs$vv, GenusPaleoIs$nn, GenusPaleoIs$tt, MM = NULL, genpars = NULL, label = "Testudinidae body size evolution mode, Holocene-Upper Miocene, genus mean CL", reset.time=TRUE)

plot(paleoGenIs)
```



```{r "Model-fitting, genera, continental" , echo=FALSE, message=FALSE, include=FALSE}

PaleoGenIsFit <- (fit3models(paleoGenIs, silent=FALSE, method="AD", pool=FALSE))
```



```{r "Model-fitting, genera, continental, table", echo=FALSE}

kable(PaleoGenIsFit, caption="Model-fitting results for testudinidae, genera, excluding insular species")

```



\newpage


## insular (excluding continental)
### individuals (insular)

```{r "paleoTS, individuals, exluding continental species", echo=FALSE, message=FALSE, fig.cap="individuals, excluding continental species"}
PPCLCon <- PleiPlioCL %>%
  dplyr::select(CL, MeanBins, Island) %>%
  dplyr::filter(CL != "NA" & Island == "y") %>%
 # dplyr::filter(PleiPlioCL$Island == "n") %>%
  mutate(tt= MeanBins) %>% # create mean age
  group_by(tt) %>% #create time bins
  summarise(mm=mean(CL), vv=var(CL), nn=n()) #create means etc. for each time bin 

PPCLCon[is.na(PPCLCon)]<-0 #subset NAs with O for n=1


sumTort <- read.csv("tortoises_summary.csv", sep=";", header=TRUE)

SumTortCon <- sumTort %>%
  dplyr::filter(Island == "y") %>%
  mutate(tt=(Mamin+Mamax)/2, vv=sdCLmm^2, nn=n, mm=meanCLmm) %>%
  dplyr::select(mm, nn, vv, tt)
 
PPESCLCon <- bind_rows(SumTortCon,PPCLCon) #ExTort, 

combiCLCon <- PPESCLCon %>%
  dplyr::filter(tt==0.0000005) %>%   #maybe find a way to automatically filter double tt values...
  mutate(nx = nn*mm) %>%
  mutate(mmall=sum(nx)/sum(nn)) %>%
  mutate(SD=sqrt(nx), d=mm-mmall) %>%
  mutate(nsd=((nx^2+d^2)*nn)) %>%
  mutate(varall=sum(nsd)/sum(nn), n=sum(nn)) %>%
  dplyr::select(mm=mmall, vv=varall, nn=n, tt) %>%
  unique() %>%
  dplyr::select(mm, nn, vv, tt)


PPCLCon <- PPESCLCon %>%
  filter(tt !=0.0000005) %>%
  bind_rows(combiCLCon)%>%
  arrange(tt)



#kable(PPCLCon, caption="paleoTS object (mm= mean CL, nn = sample size, vv = variance (CL), tt = Age)")


library(paleoTS)
paleoPPCLCon <-as.paleoTS(PPCLCon$mm, PPCLCon$vv, PPCLCon$nn, PPCLCon$tt, MM = NULL, genpars = NULL, label = "Testudinidae body size evolution mode, Holocene-Upper Miocene, CL, only insular species")
#paleoPPCL
plot(paleoPPCLCon)
```




```{r "Model-fitting, individuals, insular" , echo=FALSE, message=FALSE, include=FALSE}

PaleoPCCLConFit <- (fit3models(paleoPPCLCon, silent=FALSE, method="AD", pool=FALSE))
```



```{r "Model-fitting, individuals, insular, table", echo=FALSE}

kable(PaleoPCCLConFit, caption="Model-fitting results for testudinidae, individuals, only insular species")

```

\newpage

### species (insular)

```{r "paleoTS plot with species mean, excluding continental species", echo=FALSE, message=FALSE, fig.cap="paleoTS plot with species mean, only insular species"}

SpeciesMeanCon <- PleiPlioCL %>%
  filter(EpochBins != "Modern" & Island == "y") %>%
  group_by(EpochBins, Taxon, Age) %>%
  summarise(meanCL = mean(CL), n=n()) %>%
  merge(BINS)

SpeciesModernCon <- PleiPlioCL %>%
  filter(EpochBins == "Modern" & Island == "y") %>%
  group_by(Taxon) %>%
  summarise(meanCL=mean(CL), sdCL=sd(CL), n=n(), Age=mean(Age), EpochBins=unique(EpochBins), MeanBins=unique(MeanBins))


SpeciesModernCon[is.na(SpeciesModernCon)]<-0 #subset NAs with O for n=1



SpeciesOverviewCon <- SpeciesMeanCon %>%
#  filter(EpochBins != "Modern") %>%
  filter( !is.na(meanCL)) %>%
  group_by(EpochBins) %>%
  summarise(meanSpeciesCL = mean(meanCL), nSpecies=n(), MeanBins = mean(MeanBins)) %>%
  arrange(MeanBins) 

#kable(SpeciesOverviewCon)




SpeciesPaleoCon <- SpeciesMeanCon %>%
  dplyr::select(meanCL, MeanBins) %>%
  dplyr::filter(meanCL != "NA") %>%
  mutate(tt= MeanBins) %>% # create mean age
  group_by(tt) %>% #create time bins
  summarise(mm=mean(meanCL), vv=var(meanCL), nn=n()) #create means etc. for each time bin 

SpeciesPaleoCon[is.na(SpeciesPaleoCon)]<-0 #subset NAs with O for n=1


sumTort <- read.csv("tortoises_summary.csv", sep=";", header=TRUE)
colnames(sumTort)[6] <- "meanCL"
colnames(sumTort)[7] <- "sdCL"


sumTortoisesCon <- sumTort %>%
  dplyr::filter(Island == "y") %>%
  mutate(Taxon= as.character(Species)) %>%
  mutate(MeanBins=(Mamin+Mamax)/2) %>%
  select(Taxon, MeanBins,  meanCL, sdCL) %>% #n,
  bind_rows(SpeciesModernCon)


SumTortCon <- sumTortoisesCon %>%
  group_by(Taxon) %>%
  mutate(tt=MeanBins, vv=sdCL^2, nn=n, mm=meanCL) %>%
  dplyr::select(mm, nn, vv, tt, Taxon) %>%
  mutate(nx = nn*mm) %>%
  mutate(mmall=sum(nx)/sum(nn)) %>%
  mutate(SD=sqrt(nx), d=mm-mmall) %>%
  mutate(nsd=((nx^2+d^2)*nn)) %>%
  mutate(varall=sum(nsd)/sum(nn), n=sum(nn)) %>%
  dplyr::select(mm=mmall, vv=varall, nn=n, tt, Taxon) %>%
  unique() %>%
  dplyr::select(CL=mm, n=nn, var=vv, tt, Taxon)

# write.table(SumTort,file="SumTortModern.txt", sep="\t", row.names = FALSE) 

#boxplot(SumTort$CL, caption="Body size distribution in time bin 'modern'.")


modernMeanCon <- SumTortCon %>%
  ungroup() %>%
  select(CL, n, tt) %>%
  filter( !is.na(CL)) %>%
  group_by(tt) %>%
  summarise(mm = mean(CL), vv=var(CL), nn=n())

#bis hier alle modernen taxa zusammengefasst (summarised and MFN_testudinidae)



SpecPaleoCon <- bind_rows(modernMeanCon,SpeciesPaleoCon) %>%
  arrange(tt)


#kable(SpecPaleoCon)


paleoSpecCon <-as.paleoTS(SpecPaleoCon$mm, SpecPaleoCon$vv, SpecPaleoCon$nn, SpecPaleoCon$tt, MM = NULL, genpars = NULL, label = "Testudinidae body size evolution mode, Holocene-Upper Miocene, species mean CL, excluding contintenal species", reset.time=TRUE)

plot(paleoSpecCon)
```



```{r "Model-fitting, species, insular" , echo=FALSE, message=FALSE, include=FALSE}

PaleoSpecConFit <- (fit3models(paleoSpecCon, silent=FALSE, method="AD", pool=FALSE))
```



```{r "Model-fitting, species, insular, table", echo=FALSE}

kable(PaleoSpecIsFit, caption="Model-fitting results for testudinidae, species, only insular species")

```


\newpage


### genera (insular)

```{r "paleoTS plot with genus mean, excluding continental species", echo=FALSE, message=FALSE, fig.cap="paleoTS plot with genus mean, only insular species"}

GenusMeanCon <- PleiPlioCL %>%
  filter(EpochBins != "Modern" & Island == "y") %>%  # & EpochBins != "Upper Miocene"
  filter(!is.na(CL)) %>%
  group_by(EpochBins, Genus) %>%
  summarise(meanCL = mean(CL), n=n()) %>%
#  ungroup() %>%
#  group_by(EpochBins) %>%
  summarise(mm = mean(meanCL), nn=n(), vv=var(meanCL)) %>%
  merge(BINS) %>%
  select(mm, nn, vv, tt=MeanBins)
  
GenusMeanCon[is.na(GenusMeanCon)]<-0



GenusModernCon <- PleiPlioCL %>%
  filter(EpochBins == "Modern" & Island == "y") %>%
  filter(!is.na(CL)) %>%
  group_by(Genus) %>%
  summarise(meanCL=mean(CL), sdCL=sd(CL), n=n(), Age=mean(Age), EpochBins=unique(EpochBins), MeanBins=unique(MeanBins)) #%>%
#  summarise(mm=mean(meanCL), nn=n(), vv=var(meanCL), tt=mean(MeanBins))

GenusModernCon[is.na(GenusModernCon)]<-0


sumTort <- read.csv("tortoises_summary.csv", sep=";", header=TRUE)
colnames(sumTort)[7] <- "meanCL"
colnames(sumTort)[8] <- "sdCL"


sumTortoisesCon <- sumTort %>%
  dplyr::filter(Island == "y") %>%
  mutate(Genus= as.character(Genus)) %>%
  mutate(MeanBins=(Mamin+Mamax)/2) %>%
  select(Genus, MeanBins,  meanCL, sdCL, n) %>% #n,
  bind_rows(GenusModernIs)


SumTortCon <- sumTortoisesCon %>%
  group_by(Genus) %>%
  mutate(tt=MeanBins, vv=sdCL^2, nn=n, mm=meanCL) %>%
  dplyr::select(mm, nn, vv, tt, Genus) %>%
  mutate(nx = nn*mm) %>%
  mutate(mmall=sum(nx)/sum(nn)) %>%
  mutate(SD=sqrt(nx), d=mm-mmall) %>%
  mutate(nsd=((nx^2+d^2)*nn)) %>%
  mutate(varall=sum(nsd)/sum(nn), n=sum(nn)) %>%
  dplyr::select(mm=mmall, vv=varall, nn=n, tt, Genus) %>%
  unique() %>%
  dplyr::select(CL=mm, n=nn, var=vv, tt, Genus)

#write.table(SumTort,file="SumTortModernGenus.txt", sep="\t", row.names = FALSE)

#kable(SumTort, caption="Overview over body size means per time bin on genus level, excluding island species.")

#boxplot(SumTort$CL, caption="Body size distribution in time bin 'modern'.")


modernMeanGenusCon <- SumTortCon %>%
  ungroup() %>%
  select(CL, n, tt) %>%
  filter( !is.na(CL)) %>%
  group_by(tt) %>%
  summarise(mm = mean(CL), vv=var(CL), nn=n())

#bis hier alle modernen taxa zusammengefasst (summarised and MFN_testudinidae)


GenusPaleoCon <- modernMeanGenusCon %>%
  bind_rows(GenusMeanCon)%>%
  arrange(tt)


#kable(GenusPaleoCon)


paleoGenCon <-as.paleoTS(GenusPaleoCon$mm, GenusPaleoCon$vv, GenusPaleoCon$nn, GenusPaleoCon$tt, MM = NULL, genpars = NULL, label = "Testudinidae body size evolution mode, Holocene-Upper Miocene, genus mean CL, insular genera", reset.time=TRUE)

plot(paleoGenCon)
```



```{r "Model-fitting, genera, insular" , echo=FALSE, message=FALSE, include=FALSE}

PaleoGenConFit <- (fit3models(paleoGenCon, silent=FALSE, method="AD", pool=FALSE))
```



```{r "Model-fitting, genera, insular, table", echo=FALSE}

kable(PaleoGenConFit, caption="Model-fitting results for testudinidae, genera, only insular species")

```



\newpage

# Boxplots (continental (n) vs. Island (y) species)

```{r "Boxplot continental vs. insular, individuals", echo=FALSE, warning=FALSE, fig.cap="Boxplot continental vs. insular, individuals"}

Epoch <- as.vector(c("Modern", "Pleistocene", "Pleistocene", "Pleistocene", "Holocene",
                      "Pliocene", "Pliocene", "Miocene"))
EpochBins <- as.vector(unique(PleiPlioCL$EpochBins))
Epochs <- data.frame(Epoch,EpochBins )

Continent <- as.vector(unique(PleiPlioCL$Continent))
Con <- as.vector(c("Asia", "Africa", "Europe", "America", "America", "America", "Europe", "Europe"))
Continents <- data.frame(Continent, Con)

Island <- PleiPlioCL %>%
  dplyr::select(Taxon, CL, PL, Age, Island, Continent, EpochBins) %>%
  merge(Epochs) %>%
  merge(Continents) %>%
  dplyr::select(CL, Island, Con, Epoch, Age)
  
colnames(Island)[3] <- "Continent"

IslandBox <- Island %>%
  dplyr::select(CL, Island, Continent, Epoch, Age) %>%
  ggplot(aes(Island, CL)) + geom_boxplot() + geom_jitter(aes(colour=Continent, shape=Epoch)) +
#  facet_grid(.~Continent) +
  theme_classic() + # for white background 
#  stat_summary(fun.data = give.n, geom = "text") +
 # xlab("Sex") + ylab("Carapace length [mm]") + 
  theme(legend.background = element_rect(colour = 'black'),
        panel.border = element_rect(colour = "black", fill=NA)) +
  scale_colour_brewer(palette="Set1")

IslandBox




```



```{r "Boxplots, genera summarised", echo=FALSE, fig.cap="Boxplots continental vs. insular, species summarised"}
Island <- PleiPlioCL %>%
  dplyr::select(Taxon, CL, PL, Age, Island, Continent, EpochBins) %>%
  merge(Epochs) %>%
  merge(Continents) %>%
  dplyr::select(CL, Island, Con, Epoch, Age, Taxon)
  
colnames(Island)[3] <- "Continent"

IslandBoxSpecies <- Island %>%
  dplyr::select(CL, Island, Continent, Epoch, Age, Taxon) %>%
  group_by(Taxon, Island, Continent, Epoch) %>%
  dplyr::filter(!is.na(CL)) %>%
  summarise(meanCL=mean(CL)) %>%
  ggplot(aes(Island, meanCL)) + geom_boxplot() + geom_jitter(aes(colour=Continent, shape=Epoch)) +
#  facet_grid(.~Continent) +
  theme_classic() + # for white background 
#  stat_summary(fun.data = give.n, geom = "text") +
 # xlab("Sex") + ylab("Carapace length [mm]") + 
  theme(legend.background = element_rect(colour = 'black'),
        panel.border = element_rect(colour = "black", fill=NA)) +
  scale_colour_brewer(palette="Set1")

IslandBoxSpecies
```

```{r "Boxplot continental vs. insular, species summarised", echo=FALSE, fig.cap="Boxplot continental vs. insular, genera summarised"}
Island <- PleiPlioCL %>%
  dplyr::select(Genus, CL, PL, Age, Island, Continent, EpochBins) %>%
  merge(Epochs) %>%
  merge(Continents) %>%
  dplyr::select(CL, Island, Con, Epoch, Age, Genus)
  
colnames(Island)[3] <- "Continent"

IslandBoxGenus <- Island %>%
  dplyr::select(CL, Island, Continent, Epoch, Age, Genus) %>%
  group_by(Genus, Island, Continent, Epoch) %>%
  dplyr::filter(!is.na(CL)) %>%
  summarise(meanCL=mean(CL)) %>%
#  summarise(meanCL=mean(meanCL)) %>%
  ggplot(aes(Island, meanCL)) + geom_boxplot() + geom_jitter(aes(colour=Continent, shape=Epoch)) +
#  facet_grid(.~Continent) +
  theme_classic() + # for white background 
#  stat_summary(fun.data = give.n, geom = "text") +
 # xlab("Sex") + ylab("Carapace length [mm]") + 
  theme(legend.background = element_rect(colour = 'black'),
        panel.border = element_rect(colour = "black", fill=NA)) +
  scale_colour_brewer(palette="Set1")

IslandBoxGenus
```



```{r "Boxplots continents, insular vs. continent with age indicated",  echo=FALSE, warning=FALSE, fig.cap="Boxplots of body size (individuals) on different continents, insular vs. continent with age indicated"}
IslandBoxAge <- Island %>%
  ggplot(aes(Island, CL)) + geom_boxplot() + geom_jitter(aes(colour=Age)) +
  facet_grid(.~Continent) +
  theme_classic() + # for white background 
#  stat_summary(fun.data = give.n, geom = "text") +
 # xlab("Sex") + ylab("Carapace length [mm]") + 
  theme(legend.background = element_rect(colour = 'black'),
        panel.border = element_rect(colour = "black", fill=NA)) +
  scale_colour_gradientn(colors=c("orange", "red", "purple", "blue", "green")) #

IslandBoxAge
```
